---
title: "Analysis template"
author: "Simon Yersin"
date: '2022-03-21'
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

setwd("/Users/julian/Documents/1A_POST_DOC_LAUSANNE_2022/14_ALINE_MASTER/1A_MOUSE_EXPERIMENTS/NOVOGEN_16S/R_ANALYSIS/")

getwd()
```
#Loading library
```{r}
library(Biostrings)
library(devtools)
library(phyloseq)
library(tidyr)
library(plyr)
library(readxl)
library(dplyr)
library(stringr)
library(ggplot2)
library(vegan)
library(microbiome)
library(microViz)
library(ape)
library(ggpubr)
library(ggsignif)
library(ggprism)
library(pheatmap)
library(picante)
library(magrittr)
library(clusterCrit)
library(RColorBrewer)
library(ggcorrplot)
library(VennDiagram)
library(DESeq2)
library(ComplexHeatmap)
library(ANCOMBC)
library(tidyverse)
library(microbiomeMarker)
library(devtools)
#library(SpiecEasi)
#library(seqtime)
library(reshape2)
library(plotly)
#library(microbiomeutilities)
library(microbiome)
library(knitr)
library(data.table)
library(readr)
#library(Maaslin2)

```


#Phyloseq dataframe creation
Taxonomy table and OTU table come from the DADA2 pipeline template
## 1. Taxonomy table
```{r}
## 1. Taxonomy table
# tax contain the assigned taxonomy table
# .txt file contain sequences and corresponding assigned species
# Rank1: Kingdom, 2: Phylum, 3: Order, 4: Order, 5: Family, 6: Genus, 7:Species
tax = read.delim("taxonomy.txt", header = TRUE, sep = "\t")
#View(tax)

#Add column with ASV# for each sequence, and column with sequences from the row names
tax <- tax %>%
  mutate(ASV = paste("ASV", seq(1:nrow(tax)), sep="")) %>%
  mutate(SEQ = rownames(tax))

#View(tax)

# Assign row names as ASV number
row.names(tax) <- tax$ASV

#View(row.names(tax))

ASV_rownames <- rownames(tax)

#View(ASV_rownames)

# Check for duplicate in ASV names
which(duplicated(row.names(tax))==TRUE)

#View(tax)

# Return integrer =(0) there are no duplicates in term of all variables
# Correspondence dataframe contain first two column of tax : ASV : SEQ
correspondence <-tax[,c(8:9)]

# Change tax as matrix, remove SEQ columns
tax=as.matrix(tax)
tax <- tax[,-(8:9)]

# Create taxonomy formal Order taxonomy table for the phyloseq
taxonomy <- tax_table(tax)
``` 



## 2. OTU table
```{r}
# Importing merged sequence table, contains ASV count per samples 
#otu <- read.delim("sequence_table_JU_BA_TREATMENT_AND_CARLA.txt", header = TRUE, sep = "\t")



otu <- fread("sequence_table.txt", skip = 1)

#View(otu)


# Transposition of OTU (matrix Order) 
t_otu <- t(otu)

#First line as column name
colnames(t_otu)=t_otu[c(1),]
#Then you can erase the first row, as now it is doubled

t_otu=t_otu[-c(1),]

# Change column name
names(otu)[1] <- "ASV"

View(t_otu)


#otu_df <- as.data.frame(otu)


# OTU matrix table of 15'266 rows (sequences) and 716 columns (samples)

rownames(t_otu) <- rownames(tax)


#row.names(correspondence) <- correspondence$SEQ

# Bind together otu and correspondence 
#otu <- cbind(correspondence, otu)



# Class matrix and numerical values
t_otu<-as.matrix(t_otu)
class(t_otu) <- "numeric"

##To remove the raw form the sample name of the OTU files
#colnames(otu) <- str_remove(colnames(otu), "_raw_1.fq.gz")

#Create taxa formal Order otu_table based on otu matrix
taxa = otu_table(t_otu, taxa_are_rows = TRUE)
```

## 3. Metadata
```{r}
# Load your metadata file 
# Excel file (select excel sheet if several)
#md <- read_excel("Metadata/mystudy_metadata.xlsx", sheet = "Sheet2")
# CSV file
md <- read.csv("metadata.txt", header=TRUE, sep="\t") # verify sep which might be ; or , or other

# Remove unwanted columns
#md <- md[, -c(1:8, 11)]

# Change column name
#names(md)[3] <- "Age"

# Add other columns
#md <- md %>%
#  mutate(Lifestyle = "Traditional") %>%
#  mutate(Country = "Ethiopia")

# Modify specific data 
#md$Sex[md$Sex == "f"] <- "F"

# Reorder columns if necessary
#md <- md[,c(2,9,10, 7,8,3,4,6,5,11,1)]

# Take time to really clean and order the metadata like you want!

# If metadata from several studies, make sure the columns are in the same order, then rbind them
#md_all <- rbind(md_1, md_2, md_3)

#View(md)

# Assign row names as fastq ID (which should match the sample names/fatsq names in your OTU table)
row.names(md) <- md$fastqID

# Creating a sample data Order for phyloseq
metadata <- sample_data(md)
View(metadata)

# Save the data frame as a tab-separated text file
write.table(md, "mapMetadata.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
```

##Phyloseq creation
```{r}
# Phyloseq object creation from the sequence and taxonomy tables
physeq= phyloseq(taxa, taxonomy)

# Visualization
physeq
#15'266 taxa in 718 samples

### Metadata phyloseq object
#Addition of metadata in the phyloseq object
sample_names(physeq)

physeq = merge_phyloseq(metadata, physeq)

```

Phyloseq
- Filter OTUs, select part of the data with prune_taxa() and their counterpart for samples
- Rarefy counts (when needed) using rarefy_even_depth
- Compute alpha-diversity using estimate_richness
- Compute beta-diversity using distance
- Visualize samples using plot_ordination
- Overlay environmental variables (metadata) using envfit
- Visualize count table using plot_heatmap (useful to emphasize block structure)
- Test effect of covariate using PERMANOVA with adonis
- Find differentially abundant taxa with DESEQ2

# Phyloseq exploration
```{r}
## Can be skiped if not interesting
summarize_phyloseq(physeq)
# Visualize:
physeq 
# 15 266 taxa in 716 samples
#sample_data(physeq)
tax_table(physeq)
#otu_table(physeq)

# Convert tax_table to a data frame and add taxa names as a column
tax_df <- as.data.frame(tax_table(physeq))
tax_df$TaxaID <- rownames(tax_df)

# Reorder columns to have TaxaID first (optional)
tax_df <- tax_df[, c("TaxaID", setdiff(colnames(tax_df), "TaxaID"))]

# Write to TSV
write.table(tax_df, file = "taxonomy_table_physeq.tsv", sep = "\t", quote = FALSE, row.names = FALSE)


## accession functions
# Number of taxa in phyloseq
ntaxa(physeq)
# Number of samples in phyloseq
nsamples(physeq)
# Samples names
sample_names(physeq)
# Taxa names
taxa_names(physeq)

# Rank names = name of taxonomic levels available
rank_names(physeq)
# Name of metadata column (sample variable)
sample_variables(physeq)

# Visualize OTU table (col samples, rows ASVs) 
otu_table(physeq)[1:5,1:5]
# Visualize tax table (col taxonomy rank, rows ASVs)
tax_table(physeq)[1:5,1:5]


otu_table(physeq)[1:5,1:5]

# Total count of each sample (library size)
sample_sums(physeq)
sample_sums(physeq)[1:10]
# Total count of each taxa (overall abundance across all samples)
taxa_sums(physeq)[1:10]

# values for variable varName in sample data
get_variable(physeq, varName = "TREATMENT")
# Abundance of ASV 1:10 (or all OTU) in sample 1
get_taxa(physeq, sample_names(physeq)[1])[1:10]
# Abundance of ASV 1 in sample 1 to 10 (or all samples)
get_sample(physeq, taxa_names(physeq)[1])[1:10]


```
Processing functions
```{r}
#Filtering (taxa or samples)
#prune_taxa prunes unwanted taxa from a phyloseq object based on a vector of taxa to keep
#subset_taxa subsets unwanted taxa from a phyloseq object based on TREATMENTs that must be met

#Smoothing
# tax_glom agglomerates OTUs at a given taxonomic level. Finer taxonomic information is lost

# Rarefaction
#rarefy_even_depth downsamples all samples to the same depth and prune OTUs that disappear from all samples as a results
# transform_sample_counts applies a function to the abundance vector of each sample (can be usefull for normalization)
```




# Filtering
```{r}
# Filter the dataset to keep only relevant data
get_taxa_unique(physeq, "Kingdom")  ## Bacteria, Archaea, unassigned and Eukaryota
physeq # X taxa and Y samples
# filter out Unassigned, Chloroplast, Mitochondria and Eukaryota and others
physeq_filtered <- physeq %>%
  subset_taxa(Family != "Mitochondria") %>%
  subset_taxa(Order != "Chloroplast") %>%
  subset_taxa(Kingdom !="Unassigned") %>%
  subset_taxa(Kingdom != "Eukaryota")
physeq_filtered # new number of taxa

```

#Subset the phyloseq 
For example, several groups (countries, controls, etc)
Can be usefull if the relative abundance need to be computed separatly by groups
```{r}


# physeq for FECES
physeq_filtered_FECES <- subset_samples(physeq_filtered, SITE == "FECES")
# Removing taxa with abundance 0
physeq_filtered_FECES <- prune_taxa(taxa_sums(physeq_filtered_FECES)> 0, physeq_filtered_FECES)
# Relative abundance
physeq_filtered_FECES_ra <- transform_sample_counts(physeq_filtered_FECES, function(x) {x/sum(x)*100})
physeq_filtered_FECES # 1523 taxa and 42 samples 


otu_table(physeq_filtered_FECES_ra)

otu_table(physeq_filtered_FECES)


# FOR SMALL INTESTINE (SI)

physeq_filtered_SI <- subset_samples(physeq_filtered, SITE == "SI")
# Removing taxa with abundance 0
physeq_filtered_SI <- prune_taxa(taxa_sums(physeq_filtered_SI)> 0, physeq_filtered_SI)
# Relative abundance
physeq_filtered_SI_ra <- transform_sample_counts(physeq_filtered_SI, function(x) {x/sum(x)*100})
physeq_filtered_SI # 613 taxa and 36 samples 




```

transform_sample_counts(function(x) {x/sum(x)*100})
## Frequency plots
```{r}
# Add color palettes
# Color palette creation from Brewer
palbrew_info <- brewer.pal.info[brewer.pal.info$category == "qual", ]  # Extract color info
palbrew_all <- unlist(mapply(brewer.pal,                     # Create vector with all colors
                              palbrew_info$maxcolors,
                              rownames(palbrew_info)))
# Draw colors in pie chart
pie(rep(1, 74),                                        
    col = palbrew_all)
# Set color palette (number can be change)
pal_36 <- palbrew_all[1:36]
```



######################################

# TEST TO GET Lactobacillales, since there was no Streptococcaceae,  only 3 samples with Streptococcaceae or  Streptococcus salivarius

library(phyloseq)
library(dplyr)

# Step 1: Aggregate to Order level
physeq_order <- tax_glom(physeq, taxrank = "Order")

# Step 2: Extract taxonomy table and get taxon name(s) for Lactobacillales
tax_df <- as.data.frame(tax_table(physeq_order))
lacto_taxa <- rownames(tax_df)[tax_df$Order == "Lactobacillales"]

# Step 3: Extract OTU table and subset to Lactobacillales
otu_df <- as.data.frame(otu_table(physeq_order))
if (!taxa_are_rows(physeq_order)) {
  otu_df <- t(otu_df)
}

# Get counts for Lactobacillales
lacto_counts <- otu_df[lacto_taxa, , drop = FALSE]

# If multiple ASVs/OTUs under Lactobacillales, sum across them
lacto_counts_sum <- colSums(lacto_counts)

# Create data frame
lacto_df <- data.frame(SampleID = names(lacto_counts_sum),
                       Count = as.numeric(lacto_counts_sum))

# Step 4: Get TREATMENT metadata
meta_df <- data.frame(sample_data(physeq))
meta_df$SampleID <- rownames(meta_df)
meta_df <- meta_df[, c("SampleID", "TREATMENT")]

# Step 5: Merge counts with metadata
lacto_with_treatment <- merge(lacto_df, meta_df, by = "SampleID")

# View result
print(lacto_with_treatment)

# Step 6: Write to TSV
write.table(lacto_with_treatment, file = "Lactobacillales_counts_with_treatment.tsv",
            sep = "\t", quote = FALSE, row.names = FALSE)


###### AND THE BOXPLOT

library(phyloseq)
library(dplyr)
library(ggplot2)

# Step 1: Aggregate to Order level
physeq_order <- tax_glom(physeq, taxrank = "Order")

# Step 2: Extract taxonomy table and get taxon name(s) for Lactobacillales
tax_df <- as.data.frame(tax_table(physeq_order))
lacto_taxa <- rownames(tax_df)[tax_df$Order == "Lactobacillales"]

# Step 3: Extract OTU table and subset to Lactobacillales
otu_df <- as.data.frame(otu_table(physeq_order))
if (!taxa_are_rows(physeq_order)) {
  otu_df <- t(otu_df)
}

# Get counts for Lactobacillales
lacto_counts <- otu_df[lacto_taxa, , drop = FALSE]

# Sum across ASVs/OTUs if needed
lacto_counts_sum <- colSums(lacto_counts)

# Create data frame
lacto_df <- data.frame(SampleID = names(lacto_counts_sum),
                       Count = as.numeric(lacto_counts_sum))

# Step 4: Get metadata
meta_df <- data.frame(sample_data(physeq))
meta_df$SampleID <- rownames(meta_df)
meta_df <- meta_df[, c("SampleID", "TREATMENT")]

# Step 5: Merge
lacto_with_treatment <- merge(lacto_df, meta_df, by = "SampleID")

# Step 6: Save to TSV
write.table(lacto_with_treatment, file = "Lactobacillales_counts_with_treatment.tsv",
            sep = "\t", quote = FALSE, row.names = FALSE)

# Step 7: Boxplot
# Save boxplot as PDF
pdf("Lactobacillales_abundance_boxplot.pdf", width = 6, height = 5)

ggplot(lacto_with_treatment, aes(x = TREATMENT, y = Count, fill = TREATMENT)) +
  geom_boxplot(width = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  theme_minimal() +
  labs(title = "Lactobacillales abundance per sample",
       x = "Treatment group",
       y = "Raw abundance (counts)") +
  theme(legend.position = "none")

dev.off()


######################################


```{r}
###IF YOU HAVE CREATED SUBDATASET IN THE BEGINNING


# For FECES
physeq_filtered_FECES


# Here, change the data_subset_X according to the subdataset you want to analyze
p <- plot_richness(physeq_filtered_FECES, x = "TREATMENT", color = "TREATMENT", 
                   measures = c("Observed", "Shannon", "InvSimpson"))

# Remove default jitter points by keeping only boxplots
p <- p + 
  geom_boxplot(aes(fill = TREATMENT), outlier.shape = NA, color = "black", width = 0.6, alpha = 0.7) +
  scale_fill_brewer(palette = "Set1") +  # Better colors for fill
  scale_y_continuous(limits = c(0, NA)) +  # Remove extra space
  theme_minimal(base_size = 18) +  # Clean theme with larger text
  theme(legend.position = "right",  # Adjust legend placement
        axis.text.x = element_text(angle = 45, hjust = 1),  # Improve x-axis readability
        panel.grid.major.x = element_blank(),  # Remove unnecessary grid lines
        panel.grid.minor = element_blank()) +
  ggtitle("Alpha Diversity of physeq_filtered_FECES")

# **Explicitly remove default jitter layer**
p$layers <- p$layers[-1]  

# Display the plot in the RStudio window
print(p)

# Save the plot as a PDF file
output_filename <- "richness_physeq_filtered_FECES.pdf"

ggsave(output_filename, plot = p, width = 15, height = 10, device = "pdf")

############################### FOR SMALL INTESTINE

# For SI
physeq_filtered_SI


# Here, change the data_subset_X according to the subdataset you want to analyze
p <- plot_richness(physeq_filtered_SI, x = "TREATMENT", color = "TREATMENT", 
                   measures = c("Observed", "Shannon", "InvSimpson"))

# Remove default jitter points by keeping only boxplots
p <- p + 
  geom_boxplot(aes(fill = TREATMENT), outlier.shape = NA, color = "black", width = 0.6, alpha = 0.7) +
  scale_fill_brewer(palette = "Set1") +  # Better colors for fill
  scale_y_continuous(limits = c(0, NA)) +  # Remove extra space
  theme_minimal(base_size = 18) +  # Clean theme with larger text
  theme(legend.position = "right",  # Adjust legend placement
        axis.text.x = element_text(angle = 45, hjust = 1),  # Improve x-axis readability
        panel.grid.major.x = element_blank(),  # Remove unnecessary grid lines
        panel.grid.minor = element_blank()) +
  ggtitle("Alpha Diversity of physeq_filtered_SI")

# **Explicitly remove default jitter layer**
p$layers <- p$layers[-1]  

# Display the plot in the RStudio window
print(p)

# Save the plot as a PDF file
output_filename <- "richness_physeq_filtered_SI.pdf"

ggsave(output_filename, plot = p, width = 15, height = 10, device = "pdf")


```

###################################################################################################################

#############
# Get the relative abundance table at genus level, to check for presence of Streptococcus in samples, and also in #LYSIN and CTRL.


library(phyloseq)
library(dplyr)
library(tidyr)

# Step 1: Aggregate to Genus level
physeq_genus <- tax_glom(physeq_filtered_FECES, taxrank = "Genus")

# Step 2: Transform to relative abundance per sample
physeq_genus_ra <- transform_sample_counts(physeq_genus, function(x) x / sum(x))

# Step 3: Melt to long format for per-sample relative abundances
df_sample_ra <- psmelt(physeq_genus_ra)

# Step 4: Compute mean relative abundance per genus grouped by TREATMENT
df_treatment_ra <- df_sample_ra %>%
  group_by(TREATMENT, Genus) %>%
  summarise(Mean_RA = mean(Abundance, na.rm = TRUE)) %>%
  ungroup()

# Step 5: Write both tables to TSV files
write.table(df_sample_ra, file = "relative_abundance_by_sample_genus.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
write.table(df_treatment_ra, file = "relative_abundance_by_treatment_genus.tsv", sep = "\t", quote = FALSE, row.names = FALSE)


#############


##################################################################
### EXPLORATION OF ORDINATION PLOTS WITH ord_explore  ############
##################################################################

TEMPORARY PARAMETERS CHOSEN TO EXPLORE THE MICROBIOME COPOSITION:

Taxonomic rank : Genus
Taxa transformation compositional, proportions
Distance/Dissimilarity : bray Bray-Curtis
PCoA : Principal coordinate analysis

PCA assumes Euclidean distances and is not optimal for Bray-Curtis distances or compositional data. A better alternative would be Principal Coordinates Analysis (PCoA) or Non-metric Multidimensional Scaling (NMDS) if you're using Bray-Curtis

### No filtration:

physeq_filtered_fixed <- physeq_filtered %>%
 tax_fix(
  min_length = 4,
  unknowns = c("","Unclassified", "unclassified", "NA", "na", "n.a."),
  sep = " ", anon_unique = TRUE,
  suffix_rank = "classified"
 )


ord_explore(physeq_filtered_fixed)


# THIS FONCTION HERE WAS USED THE CREATE THE EXPLORATORY GRAPH FOR MAIER COLLABO (INTEREACTIVE TOOL) ORDINATION PLOTS




tax_fix_interactive(physeq_filtered_LYSIN_CONTROL) 

physeq_filtered_LYSIN_CONTROL <- physeq_filtered_LYSIN_CONTROL %>%
 tax_fix(
  min_length = 4,
  unknowns = c(""),
  sep = " ", anon_unique = TRUE,
  suffix_rank = "classified"
 )


ord_explore(physeq_filtered_LYSIN_CONTROL)



#SECOND for treatment analysis...

physeq_filtered_TREAT <- physeq_filtered_TREAT %>%
 tax_fix(
  min_length = 4,
  unknowns = c(""),
  sep = " ", anon_unique = TRUE,
  suffix_rank = "classified"
 )


ord_explore(physeq_filtered_TREAT)


###### TEST AVEC ALINE TREAT LYSINE VS NO SALIVARIUS


# physeq for lysin treatment comparison (1)
physeq_filtered

physeq_filtered_SIC8 <- subset_samples(physeq_filtered, SIC_ID == "SIC8A")
physeq_filtered_SIC8_NO_EB <- subset_samples(physeq_filtered_SIC8, TREATMENT != "EB")
physeq_filtered_SIC8_NO_EB_24_48 <- subset_samples(physeq_filtered_SIC8_NO_EB, TIME != 0)


physeq_filtered_SIC8
physeq_filtered_SIC8_NO_EB
physeq_filtered_SIC8_NO_EB_24_48

############################################################
############################################################
############################################################


# physeq for lysin treatment comparison only in SIC8A and SIC9A
physeq_filtered_SIC8_SIC9 <- subset_samples(physeq_filtered, SIC_ID != "SIC10A" )
physeq_filtered_SIC8_SIC9_NO_EB <- subset_samples(physeq_filtered_SIC8_SIC9, TREATMENT != "EB")
physeq_filtered_SIC8_SIC9_NO_EB_24_48 <- subset_samples(physeq_filtered_SIC8_SIC9_NO_EB, TIME != 0)

physeq_filtered_SIC8_SIC9_NO_EB_24_48


# Merge everything to the Species level
#physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES <- tax_glom(physeq_filtered_SIC8_SIC9_NO_EB_24_48, "Species", NArm = TRUE)

physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES <- tax_glom(physeq_filtered_SIC8_SIC9_NO_EB_24_48, "Species")


physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES_ra <- transform_sample_counts(physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES, function(x) {x/sum(x)*100})

physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES_ra #  141 taxa and 72 samples




otu_table(physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES_ra)





# Extract taxonomy table from the phyloseq object
tax_table_fixed <- as.data.frame(tax_table(physeq_filtered_SIC8_SIC9_NO_EB_24_48))

# Replace NA species names with their corresponding ASV names
tax_table_fixed$Species <- ifelse(is.na(tax_table_fixed$Species), rownames(tax_table_fixed), tax_table_fixed$Species)

# Update the phyloseq object with the modified taxonomy
tax_table(physeq_filtered_SIC8_SIC9_NO_EB_24_48) <- as.matrix(tax_table_fixed)

# Perform taxonomic agglomeration at the Species level
physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES <- tax_glom(physeq_filtered_SIC8_SIC9_NO_EB_24_48, "Species")

# Check if changes were applied correctly
tax_table(physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES)

physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES # 4641 taxa and 72 samples


physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES_ra <- transform_sample_counts(physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES, function(x) {x/sum(x)*100})

# I wan to use this to compare the abundance of salivarius and C. comes, and other Enterococcus
# I will also use this to re-do the analyses with only 2 SICs, excluding the SIC10 that we found to be weird
physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES_ra

############################################################
############################################################
############################################################

physeq_filtered_SIC8_SIC9_NO_EB_24_48_ra <- transform_sample_counts(physeq_filtered_SIC8_SIC9_NO_EB_24_48, function(x) {x/sum(x)*100})
physeq_filtered_SIC8_SIC9_NO_EB_24_48_ra # 8745 taxa by 7 taxonomic ranks 

otu_table(physeq_filtered_SIC8_SIC9_NO_EB_24_48_ra)


# Transform Taxa counts to relative abundance
physeq_filtered_Class_relabun <- transform_sample_counts(physeq_filtered_Class , function(OTU) OTU/sum(OTU) * 100)

#Then extract the data from the phyloseq object:
taxa_physeq_filtered_Class_relabun <- psmelt(physeq_filtered_Class_relabun)






#NOT DONE YET
physeq_filtered_TREAT_SIC8_24_48_MOD <- physeq_filtered_SIC8_NO_EB_24_48 %>%
 tax_fix(
  min_length = 4,
  unknowns = c(""),
  sep = " ", anon_unique = TRUE,
  suffix_rank = "classified"
 )


ord_explore(physeq_filtered_TREAT_SIC8_24_48_MOD)





########



### SECTION BIOMARKER DISCOVERY USING LEFSE ETC ON R (microbiomeMarker) https://yiluheihei.github.io/microbiomeMarker/articles/microbiomeMarker-vignette.html

##################
##physeq_filtered_TREAT_ra##
##################

#Five methods, LEfSe (Segata et al. 2011), metagenomeSeq (Paulson et al. 2013), ALDEx2 (Fernandes et al. 2014), ANCOM (Mandal et al. 2015), and ANCOMBC (Lin and Peddada 2020), which were developed specifically for microbiome data (contain many more zeros that RNA-seq data), are also provided in our package. All these methods have greater power to detect differentially features than simple statistical tests by incorporating more sensitive tests.

#Curently, LEfSe is the most popular tool for microbiome biomarker discovery. Here we take LEfSe method for example:




```{r}

# CONTROLS! LOOKING IF SIGNIF DIFFERENCES BETWEEN LYSIN AND CONTROLS BEFORE LYSIN TREATMENTS!!!
# IF YES, THOSE WOULD NEED TO BE EXCLUDED FROM THE SIGNIF ONES LATER AFTER TREATMENTS

# not done yet

physeq_filtered_FECES_ra_D17_D21 <- subset_samples(physeq_filtered_FECES_ra, DAY %in% c("D17", "D21"))


#Check the description...
physeq_filtered_FECES_ra_D17_D21

#Check if you managed to pick the right conditions
table(sample_data(physeq_filtered_FECES_ra_D17_D21)$TREATMENT)

#Linear discriminant analysis effect size (LEFSe)
#For all ranks, “Kingdom” “Phylum” “Class” “Order” “Family” “Genus”

#You can adjust cutoff to be more or less severe. Usually used is 2...4 is more severe... If you want all taxa level, take away the taxa_rank option.


lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_2 <- run_lefse(physeq_filtered_FECES_ra_D17_D21, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 2)

lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_3 <- run_lefse(physeq_filtered_FECES_ra_D17_D21, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 3)

lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_4 <- run_lefse(physeq_filtered_FECES_ra_D17_D21, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 4)


#Check description

lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_2


lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_3


lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_4


pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_2.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_2)
dev.off()

pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_3.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_3)
dev.off()


pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_4.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_4)
dev.off()

###### With a dotplot if you prefer!!!

# dot plot
pdf("DOT_PLOT_Lefse_biomarkers_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_2.pdf", width=10, height=5)
plot_ef_dot(lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_2)
dev.off()


###### With log2 Fold Change Graph if you're interested
# set the x axis to log2 Fold Change automatically without manual intervention

edger_physeq_filtered_FECES_ra_D17_D21_ALL <- run_edger(
    physeq_filtered_FECES_ra_D17_D21,
    group = "TREATMENT",
    pvalue_cutoff = 0.05,
    p_adjust = "fdr")


pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_FECES_ra_D17_D21_ALL_TAXA.pdf", width=10, height=5)
plot_ef_bar(edger_physeq_filtered_FECES_ra_D17_D21_ALL)
dev.off()

edger_physeq_filtered_FECES_ra_D17_D21_Genus <- run_edger(
    physeq_filtered_FECES_ra_D17_D21,
    group = "TREATMENT",
    taxa_rank = "Species",
    pvalue_cutoff = 0.05,
    p_adjust = "fdr")

pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_FECES_ra_D17_D21_rank_Species.pdf", width=10, height=5)
plot_ef_bar(edger_physeq_filtered_FECES_ra_D17_D21_Genus)
dev.off()

#####

#####



# Abundance box plot

p_abd <- plot_abundance(lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_2, group = "TREATMENT")

pdf("LEFSE_MARKERS_CUT2_Abundance_BOXPLOT_HORIZONTAL_physeq_filtered_FECES_ra_D17_D21.pdf", width=10, height=5)
p_abd
#p_abd + scale_fill_manual(values = c("SIC" = "red", "FECES" = "blue"))

dev.off()


p_abd <- plot_abundance(lef_out_physeq_filtered_FECES_ra_D17_D21_Genus_cutoff_2, group = "TREATMENT")

pdf("LEFSE_MARKERS_CUT2_Abundance_BOXPLOT_HORIZONTAL_physeq_filtered_FECES_ra_D17_D21.pdf", width=10, height=5)
p_abd
#p_abd + scale_fill_manual(values = c("SIC" = "red", "FECES" = "blue"))

dev.off()




```







```{r}

# KEEPING FECES AT DAYS 23 and 24 (48 and 72 hours post treatment)

physeq_filtered_FECES_ra_D23_D24 <- subset_samples(physeq_filtered_FECES_ra, DAY %in% c("D23", "D24"))


#Check the description...
physeq_filtered_FECES_ra_D23_D24

#Check if you managed to pick the right conditions
table(sample_data(physeq_filtered_FECES_ra_D23_D24)$TREATMENT)

#Linear discriminant analysis effect size (LEFSe)
#For all ranks, “Kingdom” “Phylum” “Class” “Order” “Family” “Genus”

#You can adjust cutoff to be more or less severe. Usually used is 2...4 is more severe... If you want all taxa level, take away the taxa_rank option.


lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2 <- run_lefse(physeq_filtered_FECES_ra_D23_D24, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 2)

lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_3 <- run_lefse(physeq_filtered_FECES_ra_D23_D24, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 3)

lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_4 <- run_lefse(physeq_filtered_FECES_ra_D23_D24, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 4)


#Check description

lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2


lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_3


lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_4


pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2)
dev.off()

pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_3.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_3)
dev.off()


pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_4.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_4)
dev.off()

###### With a dotplot if you prefer!!!

# dot plot
pdf("DOT_PLOT_Lefse_biomarkers_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2.pdf", width=10, height=5)
plot_ef_dot(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2)
dev.off()


###### With log2 Fold Change Graph if you're interested
# set the x axis to log2 Fold Change automatically without manual intervention

edger_physeq_filtered_FECES_ra_D23_D24_ALL <- run_edger(
    physeq_filtered_FECES_ra_D23_D24,
    group = "TREATMENT",
    pvalue_cutoff = 0.05,
    p_adjust = "fdr")


pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_FECES_ra_D23_D24_ALL_TAXA.pdf", width=10, height=5)
plot_ef_bar(edger_physeq_filtered_FECES_ra_D23_D24_ALL)
dev.off()

edger_physeq_filtered_FECES_ra_D23_D24_Genus <- run_edger(
    physeq_filtered_FECES_ra_D23_D24,
    group = "TREATMENT",
    taxa_rank = "Species",
    pvalue_cutoff = 0.05,
    p_adjust = "fdr")

pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_FECES_ra_D23_D24_rank_Species.pdf", width=10, height=5)
plot_ef_bar(edger_physeq_filtered_FECES_ra_D23_D24_Genus)
dev.off()

#####

#####



# Abundance box plot

p_abd <- plot_abundance(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2, group = "TREATMENT")

pdf("LEFSE_MARKERS_CUT2_Abundance_BOXPLOT_HORIZONTAL_physeq_filtered_FECES_ra_D23_D24.pdf", width=10, height=5)
p_abd
#p_abd + scale_fill_manual(values = c("SIC" = "red", "FECES" = "blue"))

dev.off()


p_abd <- plot_abundance(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2, group = "TREATMENT")

pdf("LEFSE_MARKERS_CUT2_Abundance_BOXPLOT_HORIZONTAL_physeq_filtered_FECES_ra_D23_D24.pdf", width=10, height=5)
p_abd
#p_abd + scale_fill_manual(values = c("SIC" = "red", "FECES" = "blue"))

dev.off()


####################################################


# KEEPING FECES AT DAYS D22 D23 and D24 (48 and 72 hours post treatment)

physeq_filtered_FECES_ra_D22_D23_D24 <- subset_samples(physeq_filtered_FECES_ra, DAY %in% c("D22","D23", "D24"))


#Check the description...
physeq_filtered_FECES_ra_D22_D23_D24

#Check if you managed to pick the right conditions
table(sample_data(physeq_filtered_FECES_ra_D22_D23_D24)$TREATMENT)

#Linear discriminant analysis effect size (LEFSe)
#For all ranks, “Kingdom” “Phylum” “Class” “Order” “Family” “Genus”

#You can adjust cutoff to be more or less severe. Usually used is 2...4 is more severe... If you want all taxa level, take away the taxa_rank option.


lef_out_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_2 <- run_lefse(physeq_filtered_FECES_ra_D22_D23_D24, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 2)

lef_out_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_3 <- run_lefse(physeq_filtered_FECES_ra_D22_D23_D24, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 3)

lef_out_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_4 <- run_lefse(physeq_filtered_FECES_ra_D22_D23_D24, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 4)


#Check description

lef_out_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_2


lef_out_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_3


lef_out_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_4


pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_2.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_2)
dev.off()

pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_3.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_3)
dev.off()


pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_4.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_4)
dev.off()

###### With a dotplot if you prefer!!!

# dot plot
pdf("DOT_PLOT_Lefse_biomarkers_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_2.pdf", width=10, height=5)
plot_ef_dot(lef_out_physeq_filtered_FECES_ra_D22_D23_D24_Genus_cutoff_2)
dev.off()


###### With log2 Fold Change Graph if you're interested
# set the x axis to log2 Fold Change automatically without manual intervention

edger_physeq_filtered_FECES_ra_D22_D23_D24_ALL <- run_edger(
    physeq_filtered_FECES_ra_D22_D23_D24,
    group = "TREATMENT",
    pvalue_cutoff = 0.05,
    p_adjust = "fdr")


pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_FECES_ra_D22_D23_D24_ALL_TAXA.pdf", width=10, height=5)
plot_ef_bar(edger_physeq_filtered_FECES_ra_D22_D23_D24_ALL)
dev.off()

edger_physeq_filtered_FECES_ra_D22_D23_D24_Genus <- run_edger(
    physeq_filtered_FECES_ra_D22_D23_D24,
    group = "TREATMENT",
    taxa_rank = "Species",
    pvalue_cutoff = 0.05,
    p_adjust = "fdr")

pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_FECES_ra_D22_D23_D24_rank_Species.pdf", width=10, height=5)
plot_ef_bar(edger_physeq_filtered_FECES_ra_D22_D23_D24_Genus)
dev.off()

####################################################
####################################################
# NOW for SMALL INTESTINE




# KEEPING ALL SI (only at day 24...) (72 hours post treatment)

physeq_filtered_SI


#Check the description...
physeq_filtered_SI

#Check if you managed to pick the right conditions
table(sample_data(physeq_filtered_SI)$TREATMENT)

#Linear discriminant analysis effect size (LEFSe)
#For all ranks, “Kingdom” “Phylum” “Class” “Order” “Family” “Genus”

#You can adjust cutoff to be more or less severe. Usually used is 2...4 is more severe... If you want all taxa level, take away the taxa_rank option.


lef_out_physeq_filtered_SI_Genus_cutoff_2 <- run_lefse(physeq_filtered_SI, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 2)

lef_out_physeq_filtered_SI_Genus_cutoff_3 <- run_lefse(physeq_filtered_SI, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 3)

lef_out_physeq_filtered_SI_Genus_cutoff_4 <- run_lefse(physeq_filtered_SI, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 4)


#Check description

lef_out_physeq_filtered_SI_Genus_cutoff_2


lef_out_physeq_filtered_SI_Genus_cutoff_3


lef_out_physeq_filtered_SI_Genus_cutoff_4


pdf("Lefse_biomarkers_physeq_filtered_SI_Genus_cutoff_2.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_SI_Genus_cutoff_2)
dev.off()

pdf("Lefse_biomarkers_physeq_filtered_SI_Genus_cutoff_3.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_SI_Genus_cutoff_3)
dev.off()


pdf("Lefse_biomarkers_physeq_filtered_SI_Genus_cutoff_4.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_SI_Genus_cutoff_4)
dev.off()

###### With a dotplot if you prefer!!!

# dot plot
pdf("DOT_PLOT_Lefse_biomarkers_physeq_filtered_SI_Genus_cutoff_2.pdf", width=10, height=5)
plot_ef_dot(lef_out_physeq_filtered_SI_Genus_cutoff_2)
dev.off()


###### With log2 Fold Change Graph if you're interested
# set the x axis to log2 Fold Change automatically without manual intervention

edger_physeq_filtered_SI_ALL <- run_edger(
    physeq_filtered_SI,
    group = "TREATMENT",
    pvalue_cutoff = 0.05,
    p_adjust = "fdr")


pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_SI_ALL_TAXA.pdf", width=10, height=5)
plot_ef_bar(edger_physeq_filtered_SI_ALL)
dev.off()

edger_physeq_filtered_SI_Genus <- run_edger(
    physeq_filtered_SI,
    group = "TREATMENT",
    taxa_rank = "Species",
    pvalue_cutoff = 0.05,
    p_adjust = "fdr")

pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_SI_rank_Species.pdf", width=10, height=5)
plot_ef_bar(edger_physeq_filtered_SI_Genus)
dev.off()

#####

#####



# Abundance box plot

p_abd <- plot_abundance(lef_out_physeq_filtered_SI_Genus_cutoff_2, group = "TREATMENT")

pdf("LEFSE_MARKERS_CUT2_Abundance_BOXPLOT_HORIZONTAL_physeq_filtered_SI.pdf", width=10, height=5)
p_abd
#p_abd + scale_fill_manual(values = c("SIC" = "red", "FECES" = "blue"))

dev.off()


p_abd <- plot_abundance(lef_out_physeq_filtered_SI_Genus_cutoff_2, group = "TREATMENT")

pdf("LEFSE_MARKERS_CUT2_Abundance_BOXPLOT_HORIZONTAL_physeq_filtered_SI.pdf", width=10, height=5)
p_abd
#p_abd + scale_fill_manual(values = c("SIC" = "red", "FECES" = "blue"))

dev.off()


```



```{r}

# KEEPING ALL FECES DAYS D22, D23, D24 (24, 48, 72 hours post treatment)

# not done yet

physeq_filtered_FECES_ra_D23_D24 <- subset_samples(physeq_filtered_FECES_ra, DAY %in% c("D23", "D24"))


#Check the description...
physeq_filtered_FECES_ra_D23_D24

#Check if you managed to pick the right conditions
table(sample_data(physeq_filtered_FECES_ra_D23_D24)$TREATMENT)

#Linear discriminant analysis effect size (LEFSe)
#For all ranks, “Kingdom” “Phylum” “Class” “Order” “Family” “Genus”

#You can adjust cutoff to be more or less severe. Usually used is 2...4 is more severe... If you want all taxa level, take away the taxa_rank option.


lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2 <- run_lefse(physeq_filtered_FECES_ra_D23_D24, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 2)

lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_3 <- run_lefse(physeq_filtered_FECES_ra_D23_D24, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 3)

lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_4 <- run_lefse(physeq_filtered_FECES_ra_D23_D24, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 4)


#Check description

lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2


lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_3


lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_4


pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2)
dev.off()

pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_3.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_3)
dev.off()


pdf("Lefse_biomarkers_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_4.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_4)
dev.off()

###### With a dotplot if you prefer!!!

# dot plot
pdf("DOT_PLOT_Lefse_biomarkers_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2.pdf", width=10, height=5)
plot_ef_dot(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2)
dev.off()


###### With log2 Fold Change Graph if you're interested
# set the x axis to log2 Fold Change automatically without manual intervention

edger_physeq_filtered_FECES_ra_D23_D24_ALL <- run_edger(
    physeq_filtered_FECES_ra_D23_D24,
    group = "TREATMENT",
    pvalue_cutoff = 0.05,
    p_adjust = "fdr")


pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_FECES_ra_D23_D24_ALL_TAXA.pdf", width=10, height=5)
plot_ef_bar(edger_physeq_filtered_FECES_ra_D23_D24_ALL)
dev.off()

edger_physeq_filtered_FECES_ra_D23_D24_Genus <- run_edger(
    physeq_filtered_FECES_ra_D23_D24,
    group = "TREATMENT",
    taxa_rank = "Species",
    pvalue_cutoff = 0.05,
    p_adjust = "fdr")

pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_FECES_ra_D23_D24_rank_Species.pdf", width=10, height=5)
plot_ef_bar(edger_physeq_filtered_FECES_ra_D23_D24_Genus)
dev.off()

#####

#####



# Abundance box plot

p_abd <- plot_abundance(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2, group = "TREATMENT")

pdf("LEFSE_MARKERS_CUT2_Abundance_BOXPLOT_HORIZONTAL_physeq_filtered_FECES_ra_D23_D24.pdf", width=10, height=5)
p_abd
#p_abd + scale_fill_manual(values = c("SIC" = "red", "FECES" = "blue"))

dev.off()


p_abd <- plot_abundance(lef_out_physeq_filtered_FECES_ra_D23_D24_Genus_cutoff_2, group = "TREATMENT")

pdf("LEFSE_MARKERS_CUT2_Abundance_BOXPLOT_HORIZONTAL_physeq_filtered_FECES_ra_D23_D24.pdf", width=10, height=5)
p_abd
#p_abd + scale_fill_manual(values = c("SIC" = "red", "FECES" = "blue"))

dev.off()




```





##################
##physeq_filtered_LYSIN_CONTROL_ra##
##################

#Five methods, LEfSe (Segata et al. 2011), metagenomeSeq (Paulson et al. 2013), ALDEx2 (Fernandes et al. 2014), ANCOM (Mandal et al. 2015), and ANCOMBC (Lin and Peddada 2020), which were developed specifically for microbiome data (contain many more zeros that RNA-seq data), are also provided in our package. All these methods have greater power to detect differentially features than simple statistical tests by incorporating more sensitive tests.

#Curently, LEfSe is the most popular tool for microbiome biomarker discovery. Here we take LEfSe method for example:

```{r}

# We will filter out the time point 0, because it was useful for graph visu, but there is no actual treatment in 0. Just for the 24 and 48 hours.

physeq_filtered_LYSIN_CONTROL_ra_24_48 <- subset_samples(physeq_filtered_LYSIN_CONTROL_ra, TIME %in% c("24","48"))


#Check the description...
physeq_filtered_LYSIN_CONTROL_ra_24_48

#Check if you managed to pick the right conditions
table(sample_data(physeq_filtered_LYSIN_CONTROL_ra_24_48)$TREATMENT)

#Linear discriminant analysis effect size (LEFSe)
#For all ranks, “Kingdom” “Phylum” “Class” “Order” “Family” “Genus”

#You can adjust cutoff to be more or less severe. Usually used is 2...4 is more severe... If you want all taxa level, take away the taxa_rank option.


lef_out_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_2 <- run_lefse(physeq_filtered_LYSIN_CONTROL_ra_24_48, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 2)

lef_out_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_3 <- run_lefse(physeq_filtered_LYSIN_CONTROL_ra_24_48, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 3)

lef_out_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_4 <- run_lefse(physeq_filtered_LYSIN_CONTROL_ra_24_48, wilcoxon_cutoff = 0.05, group = "TREATMENT", taxa_rank = "Genus", kw_cutoff = 0.05, norm = "CPM", multigrp_strat = TRUE, lda_cutoff = 4)


#Check description

lef_out_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_2


lef_out_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_3


lef_out_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_4


pdf("Lefse_biomarkers_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_2.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_2)
dev.off()

pdf("Lefse_biomarkers_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_3.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_3)
dev.off()


pdf("Lefse_biomarkers_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_4.pdf", width=10, height=5)
plot_ef_bar(lef_out_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_4)
dev.off()

###### With a dotplot if you prefer!!!

# dot plot
pdf("DOT_PLOT_Lefse_biomarkers_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_2.pdf", width=10, height=15)
plot_ef_dot(lef_out_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_2)
dev.off()


###### With log2 Fold Change Graph if you're interested
# set the x axis to log2 Fold Change automatically without manual intervention

edger_physeq_filtered_LYSIN_CONTROL_ra_24_48_ALL <- run_edger(
    physeq_filtered_LYSIN_CONTROL_ra_24_48,
    group = "TREATMENT",
    pvalue_cutoff = 0.05,
    p_adjust = "fdr")

edger_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus <- run_edger(
    physeq_filtered_LYSIN_CONTROL_ra_24_48,
    group = "TREATMENT",
    taxa_rank = "Genus",
    pvalue_cutoff = 0.01,
    p_adjust = "fdr")


pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_LYSIN_CONTROL_ra_24_48_ALL_TAXA.pdf", width=10, height=30)
plot_ef_bar(edger_physeq_filtered_LYSIN_CONTROL_ra_24_48_ALL)
dev.off()

pdf("Log2FoldChange_edger_dese2_analysis_physeq_filtered_LYSIN_CONTROL_ra_24_48_rank_Genus.pdf", width=10, height=30)
plot_ef_bar(edger_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus)
dev.off()

#####


#####



# Abundance box plot

p_abd <- plot_abundance(lef_out_physeq_filtered_LYSIN_CONTROL_ra_24_48_Genus_cutoff_2, group = "TREATMENT")

pdf("LEFSE_MARKERS_CUT2_Abundance_BOXPLOT_HORIZONTAL_physeq_filtered_LYSIN_CONTROL_ra_24_48.pdf", width=10, height=25)
p_abd
#p_abd + scale_fill_manual(values = c("SIC" = "red", "FECES" = "blue"))

dev.off()


####################################################
```





















#####################################################
### MAKING A BASIC HEATMAP TO SHOW CLUSTERING #######
#####################################################

pdf("HEATMAP_CLUSTERING.pdf")
physeq_filtered %>%
  tax_fix() %>%
  tax_transform("compositional", rank = "Genus") %>%
  comp_heatmap()
dev.off()

#WITH ANNOTATION....


# Load necessary libraries
library(phyloseq)
library(ComplexHeatmap)
library(viridis)

# Open a loop to iterate over different taxonomic ranks
for (rank in c("Genus", "Family", "Class", "Order", "Phylum")) {
  
  # Open PDF device for the heatmap figure
  pdf(paste0("HEATMAP_", toupper(rank), "_ANNOTATED.pdf"))
  
  # Automatically generate a color palette with distinct colors
  n_colors <- 4  # Number of colors
  legend_cols <- brewer.pal(n_colors, "Set1")  # Choose a color palette from RColorBrewer
  
  # Prepare color palette for TREATMENTs
  cols <- setNames(legend_cols, unique(sample_data(physeq_filtered)$TREATMENT))
  
  # Extract metadata
  sample_metadata <- sample_data(physeq_filtered)
  
  # Create annotation object
  sample_anno <- HeatmapAnnotation(
    TREATMENT = anno_simple(sample_metadata$TREATMENT, col = cols)
  )
  
  # Generate heatmap with annotations
  htmp <- physeq_filtered %>%
    tax_fix() %>%
    tax_transform("compositional", rank = rank) %>%
    comp_heatmap(
      top_annotation = sample_anno,
      cluster_rows = TRUE,
      colors = heat_palette(palette = viridis::turbo(15))
    )
  
  # Draw the heatmap
  draw(htmp)
  
  # Close PDF device for the heatmap figure
  dev.off()
}



#### IF YOU WANNA CHECK ALSO BY SETTING MINIMUM PREVALENCE TO MAKE GRAPH MORE LIGHT....

# Define the taxonomic ranks to iterate over
taxonomic_ranks <- c("Species", "Genus", "Family", "Class", "Order", "Phylum")

# Open a loop to iterate over different taxonomic ranks
for (rank in taxonomic_ranks) {
  
  # Define the title including rank
  title <- paste("Heatmap of Colon:", rank, "Level")
  
  # Open PDF device for the heatmap figure with appropriate title
  pdf(paste0("HEATMAP_", toupper(rank), "_ANNOTATED_MIN_PREVAL_5.pdf"), width = 15, height = 100)
  
  # Automatically generate a color palette with distinct colors
  n_colors <- 4  # Number of colors
  legend_cols <- brewer.pal(n_colors, "Set1")  # Choose a color palette from RColorBrewer
  
  # Prepare color palette for TREATMENTs
  cols <- setNames(legend_cols, unique(sample_data(physeq_filtered)$TREATMENT))
  
  # Extract metadata
  sample_metadata <- sample_data(physeq_filtered)
  
  # Create annotation object
  sample_anno <- HeatmapAnnotation(
    TREATMENT = anno_simple(sample_metadata$TREATMENT, col = cols)
  )
  
  
  # Generate heatmap with annotations
  htmp <- physeq_filtered %>%
    tax_fix(unknowns = c("faecis", "hongkongensis", "johnsonii", "massiliensis", "rubneri", "sinensis", "timonensis")) %>%
    tax_transform("compositional", rank = rank) %>%
    comp_heatmap(
      top_annotation = sample_anno,
      cluster_rows = TRUE,
      colors = heat_palette(palette = viridis::turbo(15))
    )
  
  # Draw the heatmap
  draw(htmp)
  
  # Close PDF device for the heatmap figure
  dev.off()
}


####### TO GET THE ANNOTOTATION LEGENDS BECAUSE CANT PUT IT IN THE GRAPHS!!!

# Open PDF device for annotation legends figure
pdf("ANNOTATION_LEGENDS.pdf")

# Automatically generate a color palette with distinct colors
n_colors <- 4  # Number of colors
legend_cols <- brewer.pal(n_colors, "Set1")  # Choose a color palette from RColorBrewer

# Prepare color palette for TREATMENTs
cols <- setNames(legend_cols, unique(sample_data(physeq_filtered)$TREATMENT))

# Create a plot with no data
plot.new()

# Set up the plot area
par(mar = c(1, 1, 1, 1))  # Adjust margin if needed

# Plot the legend elements
legend("topright", legend = names(cols), fill = cols, title = "TREATMENT")

# Close PDF device for annotation legends figure
dev.off()

#####################################################
#####################################################
#####################################################
#END ADVANCED HEATMAPS!!!!






####################################################


### SECTION BIOMARKER DISCOVERY USING LEFSE ETC ON R (microbiomeMarker) https://yiluheihei.github.io/microbiomeMarker/articles/microbiomeMarker-vignette.html



#Five methods, LEfSe (Segata et al. 2011), metagenomeSeq (Paulson et al. 2013), ALDEx2 (Fernandes et al. 2014), ANCOM (Mandal et al. 2015), and ANCOMBC (Lin and Peddada 2020), which were developed specifically for microbiome data (contain many more zeros that RNA-seq data), are also provided in our package. All these methods have greater power to detect differentially features than simple statistical tests by incorporating more sensitive tests.

#Curently, LEfSe is the most popular tool for microbiome biomarker discovery. Here we take LEfSe method for example:

# A TRY TOWARD AUTOMATION

# Cosse the dataset you want to work on
# Working on the colon data subset orginally named data_subset_1 (near lines 330)

# Now the names for you dataset should look like this :
#For data_subset_1 :
####################### physeq_filtered_Colon
#For data_subset_2 :
####################### physeq_filtered_Small_intestine


# Define the variables to quickly generate your different analyses and graphs
# Note : our different TREATMENTs to test are located in TREATMENT in our metadata
# Change the few variables below accordingly and run the lefse analysis
#For all ranks, “Kingdom” “Phylum” “Class” “Order” “Family” “Genus”

# NOTA BENE : if there is no marker identified, you will have to run the log2fold change part manually, independently


# Define the variables
#dataset <- physeq_filtered_Colon
#dataset_name <- deparse(substitute(physeq_filtered_Colon))
dataset <- physeq_filtered_Small_intestine
dataset_name <- deparse(substitute(physeq_filtered_Small_intestine))
control <- "DMSO"
#test <- "Fleroxacin"
test <- "Floxuridine"
taxa_level <- "Genus"
cutoff <- "1.5"



# Subset the dataset to include only the control and test TREATMENTs
physeq_lefse_TREATMENT <- subset_samples(dataset, TREATMENT %in% c(control, test))

# Check the description
print(physeq_lefse_TREATMENT)

# Check if you managed to pick the right TREATMENTs
print(table(sample_data(physeq_lefse_TREATMENT)$TREATMENT))

# Construct the object name for LEfSe results
object_name <- paste0("lefse_out_", control, "_", test, "_", taxa_level, "_cutoff_", cutoff)

# Run the LEfSe analysis on the subset dataset and assign the result to the dynamically created object name
assign(object_name, run_lefse(physeq_lefse_TREATMENT, 
                              wilcoxon_cutoff = 0.1, 
                              group = "TREATMENT", 
                              taxa_rank = taxa_level, 
                              kw_cutoff = 0.01, 
                              norm = "none",   # Adjusted normalization for relative abundances
                              multigrp_strat = FALSE,  # Simplified stratification for two-group comparison
                              lda_cutoff = as.numeric(cutoff)))

# Check the created object
print(object_name)  # This prints the name of the object
print(get(object_name))  # This prints the content of the object

# Construct the PDF file name for LEfSe bar plot
lefse_pdf_file_name <- paste0("Lefse_biomarkers_", dataset_name, "_", control, "_", test, "_", taxa_level, "_cutoff_", cutoff, ".pdf")

# Create the PDF and save the LEfSe bar plot
pdf(lefse_pdf_file_name)
plot_ef_bar(get(object_name))
dev.off()

# Construct the PDF file name for LEfSe dot plot
dotplot_pdf_file_name <- paste0("DOT_PLOT_Lefse_biomarkers_", dataset_name, "_", control, "_", test, "_", taxa_level, "_cutoff_", cutoff, ".pdf")

# Create the PDF and save the LEfSe dot plot
pdf(dotplot_pdf_file_name)
plot_ef_dot(get(object_name))
dev.off()

# Run edgeR analysis for log2 fold change
edger_physeq_lefse_TREATMENT_ALL <- run_edger(
    physeq_lefse_TREATMENT,
    group = "TREATMENT",
    pvalue_cutoff = 0.01,
    p_adjust = "fdr")

# Construct the PDF file name for log2 fold change plot
log2fc_pdf_file_name <- paste0("Log2FoldChange_edger_deseq2_pval_001_", dataset_name, "_", control, "_", test, "_ALL_TAXA.pdf")

# Create the PDF and save the log2 fold change bar plot
pdf(log2fc_pdf_file_name)
plot_ef_bar(edger_physeq_lefse_TREATMENT_ALL)
dev.off()


####################################################



####################################
## COOLER MICROVIZ STACKBAR PLOT!!##
####################################





pdf("MICROVIZ_BARPLOT_TEST_PHYLUM_TREATMENT.pdf", height = 30)

physeq_filtered %>%
  comp_barplot(
    tax_level = "Phylum", n_taxa = 8,
    bar_outline_colour = NA,
    sample_order = "bray",
    bar_width = 0.7,
    palette = distinct_palette(8, pal = "kelly"),
    taxon_renamer = toupper
  ) + coord_flip()

dev.off()






# Phylum

pdf("MICROVIZ_BARPLOT_TEST_PHYLUM_MERGED_TREATMENT.pdf", height = 30)

# Often to compare groups, average compositions are presented
phyloseq::merge_samples(physeq_filtered, group = "TREATMENT") %>%
  comp_barplot(
    tax_level = "Phylum", n_taxa = 12,
    #sample_order = c("ED", "HE", "DI"),
    bar_width = 0.8
  ) +
  coord_flip() + labs(x = NULL, y = NULL)
#> Warning: NAs introduced by coercion

dev.off()


# Class

pdf("MICROVIZ_BARPLOT_TEST_CLASS_MERGED_TREATMENT.pdf", height = 30)

# Often to compare groups, average compositions are presented
phyloseq::merge_samples(physeq_filtered, group = "TREATMENT") %>%
  comp_barplot(
    tax_level = "Class", n_taxa = 12,
    #sample_order = c("ED", "HE", "DI"),
    bar_width = 0.8
  ) +
  coord_flip() + labs(x = NULL, y = NULL)
#> Warning: NAs introduced by coercion

dev.off()





#Family

pdf("MICROVIZ_BARPLOT_TEST_FAMILY_MERGED_TREATMENT.pdf")

# Often to compare groups, average compositions are presented
phyloseq::merge_samples(physeq_filtered, group = "TREATMENT") %>% tax_fix(unknowns = c("Unknown Family")) %>% 
  comp_barplot(
    tax_level = "Family", n_taxa = 12,
    #sample_order = c("ED", "HE", "DI"),
    bar_width = 0.8
  ) +
  coord_flip() + labs(x = NULL, y = NULL)
#> Warning: NAs introduced by coercion

dev.off()




#Family grouped STRAIN

pdf("MICROVIZ_BARPLOT_TEST_FAMILY_STRAIN.pdf")

# Often to compare groups, average compositions are presented
phyloseq::merge_samples(physeq_filtered, group = "STRAIN") %>% tax_fix(unknowns = c("Unknown Family")) %>% 
  comp_barplot(
    tax_level = "Family", n_taxa = 12,
    #sample_order = c("ED", "HE", "DI"),
    bar_width = 0.8
  ) +
  coord_flip() + labs(x = NULL, y = NULL)
#> Warning: NAs introduced by coercion

dev.off()


#Family grouped STRAIN

pdf("MICROVIZ_BARPLOT_TEST_FAMILY_MERGED_STRAIN.pdf")

# Often to compare groups, average compositions are presented
phyloseq::merge_samples(physeq_filtered, group = "STRAIN") %>% tax_fix(unknowns = c("Unknown Family")) %>% 
  comp_barplot(
    tax_level = "Family", n_taxa = 12,
    #sample_order = c("ED", "HE", "DI"),
    bar_width = 0.8
  ) +
  coord_flip() + labs(x = NULL, y = NULL)
#> Warning: NAs introduced by coercion

dev.off()


# Phylum

pdf("MICROVIZ_BARPLOT_TEST_PHYLUM_MERGED_STRAIN.pdf")

# Often to compare groups, average compositions are presented
phyloseq::merge_samples(physeq_filtered, group = "STRAIN") %>%
  comp_barplot(
    tax_level = "Phylum", n_taxa = 12,
    #sample_order = c("ED", "HE", "DI"),
    bar_width = 0.8
  ) +
  coord_flip() + labs(x = NULL, y = NULL)
#> Warning: NAs introduced by coercion

dev.off()


# TEST SEPARATION OF PLOTS PHYLUM

pdf("MICROVIZ_BARPLOT_TEST_2_PHYLUM_MERGED_STRAIN.pdf")

comp_barplot(physeq_filtered,
  tax_level = "Phylum", n_taxa = 12, group_by = "STRAIN",
  sample_order = "euclidean", bar_outline_colour = NA
) %>%
  patchwork::wrap_plots(nrow = 9, guides = "collect") &
  coord_flip() & labs(x = NULL, y = NULL) &
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

dev.off()


pdf("MICROVIZ_BARPLOT_TEST_2_PHYLUM_MERGED_STRAIN.pdf")

comp_barplot(physeq_filtered,
  tax_level = "Phylum", n_taxa = 12, group_by = "STRAIN",
  sample_order = "euclidean", bar_outline_colour = NA
) %>%
  patchwork::wrap_plots(nrow = 5, guides = "collect") &
  coord_flip() & labs(x = NULL, y = NULL) &
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

dev.off()

# TEST SEPARATION OF PLOTS CLASS

pdf("MICROVIZ_BARPLOT_TEST_Class_STRAIN.pdf")

comp_barplot(physeq_filtered,
  tax_level = "Class", n_taxa = 12, group_by = "STRAIN",
  sample_order = "euclidean", bar_outline_colour = NA
) %>%
  patchwork::wrap_plots(nrow = 5, guides = "collect") &
  coord_flip() & labs(x = NULL, y = NULL) &
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

dev.off()

# TEST SEPARATION OF PLOTS GENUS

pdf("MICROVIZ_BARPLOT_TEST_Genus_STRAIN.pdf")

comp_barplot(physeq_filtered,
  tax_level = "Genus", n_taxa = 15, group_by = "STRAIN",
  sample_order = "euclidean", bar_outline_colour = NA
) %>%
  patchwork::wrap_plots(nrow = 1, guides = "collect") &
  coord_flip() & labs(x = NULL, y = NULL)
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

dev.off()


############ TES AVEC ALINE SIC INDIVIDUEL POUR VOIR TRAITEMENT LYSIN VS SANS SALIVARIUS


# TEST SEPARATION OF PLOTS GENUS







##############


pdf("MICROVIZ_BARPLOT_TEST_Genus_TREATMENT.pdf")

comp_barplot(physeq_filtered,
  tax_level = "Genus", n_taxa = 15, 
  group_by = "STRAIN",  # Keep grouping by STRAIN if needed
  facet_by = "TREATMENT",  # Add faceting by TREATMENT
  sample_order = "euclidean", 
  bar_outline_colour = NA
) %>%
  patchwork::wrap_plots(nrow = 1, guides = "collect") &
  coord_flip() & 
  labs(x = NULL, y = NULL) &
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

dev.off()



#TEST2

pdf("MICROVIZ_BARPLOT_TEST_Genus_STRAIN.pdf")

comp_barplot(physeq_filtered,
  tax_level = "Genus", n_taxa = 15, group_by = "STRAIN",
  sample_order = "euclidean", bar_outline_colour = NA,
  sample_annotation = "TREATMENT"  # Add this line to annotate by TREATMENT
) %>%
  patchwork::wrap_plots(nrow = 1, guides = "collect") &
  coord_flip() & labs(x = NULL, y = NULL) &
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

dev.off()




# TEST SEPARATION OF PLOTS Order

pdf("MICROVIZ_BARPLOT_TEST_Order_MERGED_STRAIN.pdf")

comp_barplot(physeq_filtered,
  tax_level = "Order", n_taxa = 12, group_by = "STRAIN",
  sample_order = "euclidean", bar_outline_colour = NA
) %>%
  patchwork::wrap_plots(nrow = 5, guides = "collect") &
  coord_flip() & labs(x = NULL, y = NULL) &
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

dev.off()


# TEST SEPARATION OF PLOTS FAMILY


pdf("MICROVIZ_BARPLOT_TEST_Family_MERGED_STRAIN.pdf")

physeq_filtered_prevaC <- physeq_filtered_preva %>% tax_fix(unknowns = c("Unknown Family"))

comp_barplot(physeq_filtered_prevaC,
  tax_level = "Family", n_taxa = 12, group_by = "STRAIN",
  sample_order = "euclidean", bar_outline_colour = NA
) %>%
  patchwork::wrap_plots(nrow = 5, guides = "collect") &
  coord_flip() & labs(x = NULL, y = NULL) &
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

dev.off()


# TEST SEPARATION OF PLOTS GENUS


pdf("MICROVIZ_BARPLOT_TEST_GENUS_MERGED_STRAIN.pdf")

#physeq_filtered_prevaC <- physeq_filtered_preva %>% tax_fix(unknowns = c("Unknown Family"))

comp_barplot(physeq_filtered_prevaC,
  tax_level = "Genus", n_taxa = 12, group_by = "STRAIN",
  sample_order = "euclidean", bar_outline_colour = NA
) %>%
  patchwork::wrap_plots(nrow = 5, guides = "collect") &
  coord_flip() & labs(x = NULL, y = NULL) &
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

dev.off()



###
###
###
###


# MICROVIZ CORRELATION HEATPLOT!!!!! TIME WASTED IF YOU TRY TO USE CORRELATION ON CATEGORICAL VARIABLES... 
# SO SKIP FOR NOW!!!
# Correlation is when 2 variables are quantitative...
# In our case, we would need to search for associations... not correlations....

# Calculating correlations
#By default, the cor_heatmap function will correlate all taxa to all numerical sample data, using pearson correlation method.



# PEARSON CORRELATION Phylum Level
pdf("MICROVIZ_CORRELATION_HEATPLOT_Phylum_PEARSON.pdf")

physeq_filtered_preva %>%
    #Reconding variables as numbers so they can be used in correlations
    ps_mutate(
    Sample_type_DETAILED = recode(Sample_type_DETAILED, "Chenodeoxycholic Acid_0_5" = 1, "Chenodeoxycholic Acid_0_25" = 2, "Taurochenodeoxycholic acid_1" = 3, "Taurochenodeoxycholic acid_0_5" = 4, "Taurochenodeoxycholic acid_0_25" = 5, "Lithocholic acid_0_5" = 6, "Lithocholic acid_0_25" = 7, "Taurolithocholic acid_0_25" = 8, "CTRL" = 9),
    Sample_type = recode(Sample_type, "Chenodeoxycholic Acid" = 1, "Taurochenodeoxycholic acid" = 2, "Lithocholic acid" = 3, "Taurolithocholic acid" = 4, "CTRL" = 5),
    BILE_STATES = recode(BILE_STATES, "PRIMARY" = 1, "SECONDARY" = 2, "CONJU_PRIMARY" = 3, "CONJU_SECONDARY" = 4, "CTRL" = 5),
    REGION = recode(REGION, "CPB" = 1, "HJRA" = 2),
    TREATMENT = recode(TREATMENT, "CPB352" = 1, "CPB360" = 2, "CPB396" = 3, "CBP407" = 4, "CPB477" = 5, "HJRA044" = 6, "HJRA080" = 7, "HJRA085" = 8, "HJRA179" = 9, "HJRA198" = 10),
    CONCENTRATION = recode(CONCENTRATION, "0" = 1, "0,25" = 2, "0,5" = 3, "1" = 4)) %>% 
  tax_agg("Phylum") %>%
  cor_heatmap(
    # TOP 100 selected (we have less)
    taxa = tax_top(physeq_filtered_preva, 100, by = max, rank = "Phylum"),
    cor = "pearson",
    vars = c("Sample_type_DETAILED", "Sample_type", "BILE_STATES", "REGION", "TREATMENT", "REP", "CONCENTRATION")
    #var_anno = varAnnotation(
     # Zscore = anno_var_density(fun = scale, type = "violin")
    #)
  )
dev.off()



# PEARSON CORRELATION
pdf("MICROVIZ_CORRELATION_HEATPLOT_Phylum_SPEARMAN.pdf")

physeq_filtered_preva %>%
    #Reconding variables as numbers so they can be used in correlations
    ps_mutate(
      Sample_type_DETAILED = recode(Sample_type_DETAILED, "Chenodeoxycholic Acid_0_5" = 1, "Chenodeoxycholic Acid_0_25" = 2, "Taurochenodeoxycholic acid_1" = 3, "Taurochenodeoxycholic acid_0_5" = 4, "Taurochenodeoxycholic acid_0_25" = 5, "Lithocholic acid_0_5" = 6, "Lithocholic acid_0_25" = 7, "Taurolithocholic acid_0_25" = 8, "CTRL" = 9),
    Sample_type = recode(Sample_type, "Chenodeoxycholic Acid" = 1, "Taurochenodeoxycholic acid" = 2, "Lithocholic acid" = 3, "Taurolithocholic acid" = 4, "CTRL" = 5),
    BILE_STATES = recode(BILE_STATES, "PRIMARY" = 1, "SECONDARY" = 2, "CONJU_PRIMARY" = 3, "CONJU_SECONDARY" = 4, "CTRL" = 5),
    REGION = recode(REGION, "CPB" = 1, "HJRA" = 2),
    TREATMENT = recode(TREATMENT, "CPB352" = 1, "CPB360" = 2, "CPB396" = 3, "CBP407" = 4, "CPB477" = 5, "HJRA044" = 6, "HJRA080" = 7, "HJRA085" = 8, "HJRA179" = 9, "HJRA198" = 10),
    CONCENTRATION = recode(CONCENTRATION, "0" = 1, "0,25" = 2, "0,5" = 3, "1" = 4)) %>% 
  tax_agg("Phylum") %>%
  cor_heatmap(
    # TOP 100 selected (we have less)
    taxa = tax_top(physeq_filtered_preva, 100, by = max, rank = "Phylum"),
    cor = "spearman",
    vars = c("Sample_type_DETAILED", "Sample_type", "BILE_STATES", "REGION", "TREATMENT", "REP", "CONCENTRATION")
    #var_anno = varAnnotation(
    #  Zscore = anno_var_density(fun = scale, type = "violin")
    #)
  )
dev.off()




# PEARSON CORRELATION Order Level
pdf("MICROVIZ_CORRELATION_HEATPLOT_Order_PEARSON.pdf")

physeq_filtered_preva %>%
    #Reconding variables as numbers so they can be used in correlations
    ps_mutate(
      Sample_type_DETAILED = recode(Sample_type_DETAILED, "Chenodeoxycholic Acid_0_5" = 1, "Chenodeoxycholic Acid_0_25" = 2, "Taurochenodeoxycholic acid_1" = 3, "Taurochenodeoxycholic acid_0_5" = 4, "Taurochenodeoxycholic acid_0_25" = 5, "Lithocholic acid_0_5" = 6, "Lithocholic acid_0_25" = 7, "Taurolithocholic acid_0_25" = 8, "CTRL" = 9),
    Sample_type = recode(Sample_type, "Chenodeoxycholic Acid" = 1, "Taurochenodeoxycholic acid" = 2, "Lithocholic acid" = 3, "Taurolithocholic acid" = 4, "CTRL" = 5),
    BILE_STATES = recode(BILE_STATES, "PRIMARY" = 1, "SECONDARY" = 2, "CONJU_PRIMARY" = 3, "CONJU_SECONDARY" = 4, "CTRL" = 5),
    REGION = recode(REGION, "CPB" = 1, "HJRA" = 2),
    TREATMENT = recode(TREATMENT, "CPB352" = 1, "CPB360" = 2, "CPB396" = 3, "CBP407" = 4, "CPB477" = 5, "HJRA044" = 6, "HJRA080" = 7, "HJRA085" = 8, "HJRA179" = 9, "HJRA198" = 10),
    CONCENTRATION = recode(CONCENTRATION, "0" = 1, "0,25" = 2, "0,5" = 3, "1" = 4)) %>% 
  tax_agg("Order") %>%
  cor_heatmap(
    # TOP 100 selected (we have less)
    taxa = tax_top(physeq_filtered_preva, 100, by = max, rank = "Order"),
    cor = "pearson",
    vars = c("Sample_type_DETAILED", "Sample_type", "BILE_STATES", "REGION", "TREATMENT", "REP", "CONCENTRATION")
    #var_anno = varAnnotation(
     # Zscore = anno_var_density(fun = scale, type = "violin")
    #)
  )
dev.off()



# PEARSON CORRELATION
pdf("MICROVIZ_CORRELATION_HEATPLOT_Order_SPEARMAN.pdf")

physeq_filtered_preva %>%
    #Reconding variables as numbers so they can be used in correlations
    ps_mutate(
      Sample_type_DETAILED = recode(Sample_type_DETAILED, "Chenodeoxycholic Acid_0_5" = 1, "Chenodeoxycholic Acid_0_25" = 2, "Taurochenodeoxycholic acid_1" = 3, "Taurochenodeoxycholic acid_0_5" = 4, "Taurochenodeoxycholic acid_0_25" = 5, "Lithocholic acid_0_5" = 6, "Lithocholic acid_0_25" = 7, "Taurolithocholic acid_0_25" = 8, "CTRL" = 9),
    Sample_type = recode(Sample_type, "Chenodeoxycholic Acid" = 1, "Taurochenodeoxycholic acid" = 2, "Lithocholic acid" = 3, "Taurolithocholic acid" = 4, "CTRL" = 5),
    BILE_STATES = recode(BILE_STATES, "PRIMARY" = 1, "SECONDARY" = 2, "CONJU_PRIMARY" = 3, "CONJU_SECONDARY" = 4, "CTRL" = 5),
    REGION = recode(REGION, "CPB" = 1, "HJRA" = 2),
    TREATMENT = recode(TREATMENT, "CPB352" = 1, "CPB360" = 2, "CPB396" = 3, "CBP407" = 4, "CPB477" = 5, "HJRA044" = 6, "HJRA080" = 7, "HJRA085" = 8, "HJRA179" = 9, "HJRA198" = 10),
    CONCENTRATION = recode(CONCENTRATION, "0" = 1, "0,25" = 2, "0,5" = 3, "1" = 4)) %>% 
  tax_agg("Order") %>%
  cor_heatmap(
    # TOP 100 selected (we have less)
    taxa = tax_top(physeq_filtered_preva, 100, by = max, rank = "Order"),
    cor = "spearman",
    vars = c("Sample_type_DETAILED", "Sample_type", "BILE_STATES", "REGION", "TREATMENT", "REP", "CONCENTRATION")
    #var_anno = varAnnotation(
    #  Zscore = anno_var_density(fun = scale, type = "violin")
    #)
  )
dev.off()






# PEARSON CORRELATION Class Level
pdf("MICROVIZ_CORRELATION_HEATPLOT_Class_PEARSON.pdf")

physeq_filtered_preva %>%
    #Reconding variables as numbers so they can be used in correlations
    ps_mutate(
      Sample_type_DETAILED = recode(Sample_type_DETAILED, "Chenodeoxycholic Acid_0_5" = 1, "Chenodeoxycholic Acid_0_25" = 2, "Taurochenodeoxycholic acid_1" = 3, "Taurochenodeoxycholic acid_0_5" = 4, "Taurochenodeoxycholic acid_0_25" = 5, "Lithocholic acid_0_5" = 6, "Lithocholic acid_0_25" = 7, "Taurolithocholic acid_0_25" = 8, "CTRL" = 9),
    Sample_type = recode(Sample_type, "Chenodeoxycholic Acid" = 1, "Taurochenodeoxycholic acid" = 2, "Lithocholic acid" = 3, "Taurolithocholic acid" = 4, "CTRL" = 5),
    BILE_STATES = recode(BILE_STATES, "PRIMARY" = 1, "SECONDARY" = 2, "CONJU_PRIMARY" = 3, "CONJU_SECONDARY" = 4, "CTRL" = 5),
    REGION = recode(REGION, "CPB" = 1, "HJRA" = 2),
    TREATMENT = recode(TREATMENT, "CPB352" = 1, "CPB360" = 2, "CPB396" = 3, "CBP407" = 4, "CPB477" = 5, "HJRA044" = 6, "HJRA080" = 7, "HJRA085" = 8, "HJRA179" = 9, "HJRA198" = 10),
    CONCENTRATION = recode(CONCENTRATION, "0" = 1, "0,25" = 2, "0,5" = 3, "1" = 4)) %>% 
  tax_agg("Class") %>%
  cor_heatmap(
    # TOP 100 selected (we have less)
    taxa = tax_top(physeq_filtered_preva, 100, by = max, rank = "Class"),
    cor = "pearson",
    vars = c("Sample_type_DETAILED", "Sample_type", "BILE_STATES", "REGION", "TREATMENT", "REP", "CONCENTRATION")
    #var_anno = varAnnotation(
     # Zscore = anno_var_density(fun = scale, type = "violin")
    #)
  )
dev.off()



# PEARSON CORRELATION
pdf("MICROVIZ_CORRELATION_HEATPLOT_Class_SPEARMAN.pdf")

physeq_filtered_preva %>%
    #Reconding variables as numbers so they can be used in correlations
    ps_mutate(
      Sample_type_DETAILED = recode(Sample_type_DETAILED, "Chenodeoxycholic Acid_0_5" = 1, "Chenodeoxycholic Acid_0_25" = 2, "Taurochenodeoxycholic acid_1" = 3, "Taurochenodeoxycholic acid_0_5" = 4, "Taurochenodeoxycholic acid_0_25" = 5, "Lithocholic acid_0_5" = 6, "Lithocholic acid_0_25" = 7, "Taurolithocholic acid_0_25" = 8, "CTRL" = 9),
    Sample_type = recode(Sample_type, "Chenodeoxycholic Acid" = 1, "Taurochenodeoxycholic acid" = 2, "Lithocholic acid" = 3, "Taurolithocholic acid" = 4, "CTRL" = 5),
    BILE_STATES = recode(BILE_STATES, "PRIMARY" = 1, "SECONDARY" = 2, "CONJU_PRIMARY" = 3, "CONJU_SECONDARY" = 4, "CTRL" = 5),
    REGION = recode(REGION, "CPB" = 1, "HJRA" = 2),
    TREATMENT = recode(TREATMENT, "CPB352" = 1, "CPB360" = 2, "CPB396" = 3, "CBP407" = 4, "CPB477" = 5, "HJRA044" = 6, "HJRA080" = 7, "HJRA085" = 8, "HJRA179" = 9, "HJRA198" = 10),
    CONCENTRATION = recode(CONCENTRATION, "0" = 1, "0,25" = 2, "0,5" = 3, "1" = 4)) %>% 
  tax_agg("Class") %>%
  cor_heatmap(
    # TOP 100 selected (we have less)
    taxa = tax_top(physeq_filtered_preva, 100, by = max, rank = "Class"),
    cor = "spearman",
    vars = c("Sample_type_DETAILED", "Sample_type", "BILE_STATES", "REGION", "TREATMENT", "REP", "CONCENTRATION")
    #var_anno = varAnnotation(
    #  Zscore = anno_var_density(fun = scale, type = "violin")
    #)
  )
dev.off()







# PEARSON CORRELATION FAMILY LEVEL
pdf("MICROVIZ_CORRELATION_HEATPLOT_FAMILY_PEARSON.pdf")

physeq_filtered_preva %>%
    #Reconding variables as numbers so they can be used in correlations
    ps_mutate(
    Sample_type_DETAILED = recode(Sample_type_DETAILED, "Chenodeoxycholic Acid_0_5" = 1, "Chenodeoxycholic Acid_0_25" = 2, "Taurochenodeoxycholic acid_1" = 3, "Taurochenodeoxycholic acid_0_5" = 4, "Taurochenodeoxycholic acid_0_25" = 5, "Lithocholic acid_0_5" = 6, "Lithocholic acid_0_25" = 7, "Taurolithocholic acid_0_25" = 8, "CTRL" = 9),
    Sample_type = recode(Sample_type, "Chenodeoxycholic Acid" = 1, "Taurochenodeoxycholic acid" = 2, "Lithocholic acid" = 3, "Taurolithocholic acid" = 4, "CTRL" = 5),
    BILE_STATES = recode(BILE_STATES, "PRIMARY" = 1, "SECONDARY" = 2, "CONJU_PRIMARY" = 3, "CONJU_SECONDARY" = 4, "CTRL" = 5),
    REGION = recode(REGION, "CPB" = 1, "HJRA" = 2),
    TREATMENT = recode(TREATMENT, "CPB352" = 1, "CPB360" = 2, "CPB396" = 3, "CBP407" = 4, "CPB477" = 5, "HJRA044" = 6, "HJRA080" = 7, "HJRA085" = 8, "HJRA179" = 9, "HJRA198" = 10),
    CONCENTRATION = recode(CONCENTRATION, "0" = 1, "0,25" = 2, "0,5" = 3, "1" = 4)) %>% 
  tax_agg("Family") %>%
  cor_heatmap(
    # TOP 100 selected (we have less)
    taxa = tax_top(physeq_filtered_preva, 100, by = max, rank = "Family"),
    cor = "pearson",
    vars = c("Sample_type_DETAILED", "Sample_type", "BILE_STATES", "REGION", "TREATMENT", "REP", "CONCENTRATION")
    #var_anno = varAnnotation(
     # Zscore = anno_var_density(fun = scale, type = "violin")
    #)
  )
dev.off()



# PEARSON CORRELATION
pdf("MICROVIZ_CORRELATION_HEATPLOT_FAMILY_SPEARMAN.pdf")

physeq_filtered_preva %>%
    #Reconding variables as numbers so they can be used in correlations
    ps_mutate(
    Sample_type_DETAILED = recode(Sample_type_DETAILED, "Chenodeoxycholic Acid_0_5" = 1, "Chenodeoxycholic Acid_0_25" = 2, "Taurochenodeoxycholic acid_1" = 3, "Taurochenodeoxycholic acid_0_5" = 4, "Taurochenodeoxycholic acid_0_25" = 5, "Lithocholic acid_0_5" = 6, "Lithocholic acid_0_25" = 7, "Taurolithocholic acid_0_25" = 8, "CTRL" = 9),
    Sample_type = recode(Sample_type, "Chenodeoxycholic Acid" = 1, "Taurochenodeoxycholic acid" = 2, "Lithocholic acid" = 3, "Taurolithocholic acid" = 4, "CTRL" = 5),
    BILE_STATES = recode(BILE_STATES, "PRIMARY" = 1, "SECONDARY" = 2, "CONJU_PRIMARY" = 3, "CONJU_SECONDARY" = 4, "CTRL" = 5),
    REGION = recode(REGION, "CPB" = 1, "HJRA" = 2),
    TREATMENT = recode(TREATMENT, "CPB352" = 1, "CPB360" = 2, "CPB396" = 3, "CBP407" = 4, "CPB477" = 5, "HJRA044" = 6, "HJRA080" = 7, "HJRA085" = 8, "HJRA179" = 9, "HJRA198" = 10),
    CONCENTRATION = recode(CONCENTRATION, "0" = 1, "0,25" = 2, "0,5" = 3, "1" = 4)) %>% 
  tax_agg("Family") %>%
  cor_heatmap(
    # TOP 100 selected (we have less)
    taxa = tax_top(physeq_filtered_preva, 100, by = max, rank = "Family"),
    cor = "spearman",
    vars = c("Sample_type_DETAILED", "Sample_type", "BILE_STATES", "REGION", "TREATMENT", "REP", "CONCENTRATION")
    #var_anno = varAnnotation(
    #  Zscore = anno_var_density(fun = scale, type = "violin")
    #)
  )
dev.off()





# PEARSON CORRELATION Genus Level
pdf("MICROVIZ_CORRELATION_HEATPLOT_Genus_PEARSON.pdf")

physeq_filtered_preva %>%
    #Reconding variables as numbers so they can be used in correlations
    ps_mutate(
    Sample_type_DETAILED = recode(Sample_type_DETAILED, "Chenodeoxycholic Acid_0_5" = 1, "Chenodeoxycholic Acid_0_25" = 2, "Taurochenodeoxycholic acid_1" = 3, "Taurochenodeoxycholic acid_0_5" = 4, "Taurochenodeoxycholic acid_0_25" = 5, "Lithocholic acid_0_5" = 6, "Lithocholic acid_0_25" = 7, "Taurolithocholic acid_0_25" = 8, "CTRL" = 9),
    Sample_type = recode(Sample_type, "Chenodeoxycholic Acid" = 1, "Taurochenodeoxycholic acid" = 2, "Lithocholic acid" = 3, "Taurolithocholic acid" = 4, "CTRL" = 5),
    BILE_STATES = recode(BILE_STATES, "PRIMARY" = 1, "SECONDARY" = 2, "CONJU_PRIMARY" = 3, "CONJU_SECONDARY" = 4, "CTRL" = 5),
    REGION = recode(REGION, "CPB" = 1, "HJRA" = 2),
    TREATMENT = recode(TREATMENT, "CPB352" = 1, "CPB360" = 2, "CPB396" = 3, "CBP407" = 4, "CPB477" = 5, "HJRA044" = 6, "HJRA080" = 7, "HJRA085" = 8, "HJRA179" = 9, "HJRA198" = 10),
    CONCENTRATION = recode(CONCENTRATION, "0" = 1, "0,25" = 2, "0,5" = 3, "1" = 4)) %>% 
  tax_agg("Genus") %>%
  cor_heatmap(
    # TOP 100 selected (we have less)
    taxa = tax_top(physeq_filtered_preva, 100, by = max, rank = "Genus"),
    cor = "pearson",
    vars = c("Sample_type_DETAILED", "Sample_type", "BILE_STATES", "REGION", "TREATMENT", "REP", "CONCENTRATION")
    #var_anno = varAnnotation(
     # Zscore = anno_var_density(fun = scale, type = "violin")
    #)
  )
dev.off()



# PEARSON CORRELATION
pdf("MICROVIZ_CORRELATION_HEATPLOT_Genus_SPEARMAN.pdf")

physeq_filtered_preva %>%
    #Reconding variables as numbers so they can be used in correlations
    ps_mutate(
    Sample_type_DETAILED = recode(Sample_type_DETAILED, "Chenodeoxycholic Acid_0_5" = 1, "Chenodeoxycholic Acid_0_25" = 2, "Taurochenodeoxycholic acid_1" = 3, "Taurochenodeoxycholic acid_0_5" = 4, "Taurochenodeoxycholic acid_0_25" = 5, "Lithocholic acid_0_5" = 6, "Lithocholic acid_0_25" = 7, "Taurolithocholic acid_0_25" = 8, "CTRL" = 9),
    Sample_type = recode(Sample_type, "Chenodeoxycholic Acid" = 1, "Taurochenodeoxycholic acid" = 2, "Lithocholic acid" = 3, "Taurolithocholic acid" = 4, "CTRL" = 5),
    BILE_STATES = recode(BILE_STATES, "PRIMARY" = 1, "SECONDARY" = 2, "CONJU_PRIMARY" = 3, "CONJU_SECONDARY" = 4, "CTRL" = 5),
    REGION = recode(REGION, "CPB" = 1, "HJRA" = 2),
    TREATMENT = recode(TREATMENT, "CPB352" = 1, "CPB360" = 2, "CPB396" = 3, "CBP407" = 4, "CPB477" = 5, "HJRA044" = 6, "HJRA080" = 7, "HJRA085" = 8, "HJRA179" = 9, "HJRA198" = 10),
    CONCENTRATION = recode(CONCENTRATION, "0" = 1, "0,25" = 2, "0,5" = 3, "1" = 4)) %>% 
  tax_agg("Genus") %>%
  cor_heatmap(
    # TOP 100 selected (we have less)
    taxa = tax_top(physeq_filtered_preva, 100, by = max, rank = "Genus"),
    cor = "spearman",
    vars = c("Sample_type_DETAILED", "Sample_type", "BILE_STATES", "REGION", "TREATMENT", "REP", "CONCENTRATION")
    #var_anno = varAnnotation(
    #  Zscore = anno_var_density(fun = scale, type = "violin")
    #)
  )
dev.off()




#END MICROVIZ CORRELATION HEATPLOT!!!!!






# Microviz ORDINATION PLOTS


#PRE-FILTERING RARE TAXA BY PREVALENCE....

physeq_filtered_preva <- physeq_filtered %>%
  tax_fix() %>%
  tax_filter(min_prevalence = 2, tax_level = NA)


physeq_filtered_preva

physeq_filtered


pdf("MICROVIZ_ORDINATION_PHYLUM_BRAY_PCoA.pdf")

physeq_filtered_preva %>%
 tax_transform(rank = "Phylum", trans = "compositional") %>%
 dist_calc(dist = "bray") %>%
 ord_calc(
  method = "PCoA"
 ) %>% 
 ord_plot(
  axes = c(1, 2),
  colour = "Sample_type", fill = "Sample_type",
  shape = "circle", alpha = 0.5,
  size = 2
 ) +
 ggplot2::stat_ellipse(
  ggplot2::aes(colour = Sample_type)
 )

dev.off()




pdf("MICROVIZ_ORDINATION_PHYLUM_BRAY_PCoA_DETAILED.pdf")

physeq_filtered_preva %>%
 tax_transform(rank = "Phylum", trans = "compositional") %>%
 dist_calc(dist = "bray") %>%
 ord_calc(
  method = "PCoA"
 ) %>% 
 ord_plot(
  axes = c(1, 2),
  colour = "Sample_type_DETAILED", fill = "Sample_type_DETAILED",
  shape = "circle", alpha = 0.5,
  size = 2
 ) +
 ggplot2::stat_ellipse(
  ggplot2::aes(colour = Sample_type_DETAILED)
 )

dev.off()



pdf("MICROVIZ_ORDINATION_Class_BRAY_PCoA_DETAILED.pdf")

physeq_filtered_preva %>% #tax_fix(unknowns = c("Unknown Family"))
 tax_transform(rank = "Class", trans = "compositional") %>%
 dist_calc(dist = "bray") %>%
 ord_calc(
  method = "PCoA"
 ) %>% 
 ord_plot(
  axes = c(1, 2),
  colour = "Sample_type_DETAILED", fill = "Sample_type_DETAILED",
  shape = "circle", alpha = 0.5,
  size = 2
 ) +
 ggplot2::stat_ellipse(
  ggplot2::aes(colour = Sample_type_DETAILED)
 )

dev.off()


pdf("MICROVIZ_ORDINATION_Family_BRAY_PCoA_DETAILED.pdf")

physeq_filtered_preva %>% #tax_fix(unknowns = c("Unknown Family"))
 tax_transform(rank = "Family", trans = "compositional") %>%
 dist_calc(dist = "bray") %>%
 ord_calc(
  method = "PCoA"
 ) %>% 
 ord_plot(
  axes = c(1, 2),
  colour = "Sample_type_DETAILED", fill = "Sample_type_DETAILED",
  shape = "circle", alpha = 0.5,
  size = 2
 ) +
 ggplot2::stat_ellipse(
  ggplot2::aes(colour = Sample_type_DETAILED)
 )

dev.off()



pdf("MICROVIZ_ORDINATION_Genus_BRAY_PCoA_DETAILED.pdf")

physeq_filtered_preva %>% #tax_fix(unknowns = c("Unknown Family"))
 tax_transform(rank = "Genus", trans = "compositional") %>%
 dist_calc(dist = "bray") %>%
 ord_calc(
  method = "PCoA"
 ) %>% 
 ord_plot(
  axes = c(1, 2),
  colour = "Sample_type_DETAILED", fill = "Sample_type_DETAILED",
  shape = "circle", alpha = 0.5,
  size = 2
 ) +
 ggplot2::stat_ellipse(
  ggplot2::aes(colour = Sample_type_DETAILED)
 )

dev.off()



pdf("MICROVIZ_ORDINATION_Genus_BRAY_PCoA_DETAILED_NOURIS_STATE.pdf")

physeq_filtered_preva %>% #tax_fix(unknowns = c("Unknown Family"))
 tax_transform(rank = "Genus", trans = "compositional") %>%
 dist_calc(dist = "bray") %>%
 ord_calc(
  method = "PCoA"
 ) %>% 
 ord_plot(
  axes = c(1, 2),
  colour = "Sample_type_DETAILED", fill = "Sample_type_DETAILED",
  shape = "circle", alpha = 0.5,
  size = 2
 ) +
 ggplot2::stat_ellipse(
  ggplot2::aes(colour = NUT_STAT)
 )

dev.off()




#NOW WORKING WITH NEW METADATA
pdf("MICROVIZ_ORDINATION_phylum_BRAY_PCoA_NUT_STAT.pdf")

physeq_filtered %>% #tax_fix() %>% tax_fix(unknowns = c("Incertae Sedis", "Unknown Family Family")) %>% tax_fix(unknowns = c("Unknown Family Family")) %>%
 tax_transform(rank = "Phylum", trans = "compositional") %>%
 dist_calc(dist = "bray") %>%
 ord_calc(
  method = "PCoA"
 ) %>% 
 ord_plot(
  axes = c(1, 2),
  colour = "NUT_STAT", fill = "NUT_STAT",
  shape = "circle", alpha = 0.5,
  size = 2
 ) +
 ggplot2::stat_ellipse(
  ggplot2::aes(colour = Sample_type_DETAILED)
 )

dev.off()



# OTHER VERSION
pdf("MICROVIZ_ORDINATION_phylum_BRAY_PCoA_NUT_STAT_ONLY.pdf", width=12, height = 12)

physeq_filtered %>% #tax_fix() %>% tax_fix(unknowns = c("Incertae Sedis", "Unknown Family Family")) %>% tax_fix(unknowns = c("Unknown Family Family")) %>%
 tax_transform(rank = "Phylum", trans = "compositional") %>%
 dist_calc(dist = "bray") %>%
 ord_calc(
  method = "PCoA"
 ) %>% 
 ord_plot(
  axes = c(1, 2),
  colour = "NUT_STAT_DETAILED", fill = "NUT_STAT_DETAILED",
  shape = "NUT_STAT", alpha = 0.5,
  size = 2
 ) #+
# ggplot2::stat_ellipse(
#  ggplot2::aes(colour = Sample_type)
# )

dev.off()

# OTHER VERSION WITH ELLIPSES
pdf("MICROVIZ_ORDINATION_phylum_BRAY_PCoA_NUT_STAT_ONLY_ELLIPSE.pdf", width=12, height = 12)

physeq_filtered %>% #tax_fix() %>% tax_fix(unknowns = c("Incertae Sedis", "Unknown Family Family")) %>% tax_fix(unknowns = c("Unknown Family Family")) %>%
 tax_transform(rank = "Phylum", trans = "compositional") %>%
 dist_calc(dist = "bray") %>%
 ord_calc(
  method = "PCoA"
 ) %>% 
 ord_plot(
  axes = c(1, 2),
  colour = "Sample_type", fill = "Sample_type",
  shape = "NUT_STAT_DETAILED", alpha = 0.5,
  size = 2
 ) +
 ggplot2::stat_ellipse(
  ggplot2::aes(colour = Sample_type)
 )

dev.off()




### YET ANOTHER VERSION
### THIS ONE PARRTICULARILY INTERESTING FOR OUR BILE ACID ANALYSES (Phylum)

pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_PHYLUM.pdf", width=100, height = 500)

taxa_physeq_filtered_phylum_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Phylum Relative Abundance") +
  facet_wrap(NUT_STAT~Phylum, scales= "free_x", ncol=10) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


### THIS ONE PARRTICULARILY INTERESTING FOR OUR BILE ACID ANALYSES (Class)

# Merge everything to the Class level
physeq_filtered_Class <- tax_glom(physeq_filtered, "Class", NArm = TRUE)

#View(physeq_filtered_Class) 

# Transform Taxa counts to relative abundance
physeq_filtered_Class_relabun <- transform_sample_counts(physeq_filtered_Class , function(OTU) OTU/sum(OTU) * 100)

#Then extract the data from the phyloseq object:
taxa_physeq_filtered_Class_relabun <- psmelt(physeq_filtered_Class_relabun)



pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Class.pdf", width=30, height = 50)

taxa_physeq_filtered_Class_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Class Relative Abundance") +
  facet_wrap(TREATMENT~Class, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Class_less_than_5PERC.pdf", width=30, height = 50)

taxa_physeq_filtered_Class_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Class Relative Abundance") +
  ylim(0, 5) +
  facet_wrap(TREATMENT~Class, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Class_less_than_0dot5_PERC.pdf", width=30, height = 50)

taxa_physeq_filtered_Class_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Class Relative Abundance") +
  ylim(0, 0.5) +
  facet_wrap(TREATMENT~Class, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Class_less_than_005_PERC.pdf", width=30, height = 50)

taxa_physeq_filtered_Class_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Class Relative Abundance") +
  ylim(0, 0.005) +
  facet_wrap(TREATMENT~Class, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()



# make a baTREATMENT heatmap
#pdf("BATREATMENT_PRESENCE_ABSENCE_HEATMAP_ALL_SAMPLES.pdf", width = 20, height=20)
pdf("BATREATMENT_HEATMAP_Class.pdf")

phyloseq::plot_heatmap(physeq_filtered_Class, "PCoA", "bray", taxa.label="Class", low="Red", high="Blue", na.value="white", title = "Class Level Heatmap PCoA-Bray") +
theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5))
dev.off()




### THIS ONE PARRTICULARILY INTERESTING FOR OUR BILE ACID ANALYSES (Family)


# Merge everything to the Family level
physeq_filtered_Family <- tax_glom(physeq_filtered, "Family", NArm = TRUE)

#View(physeq_filtered_Family) 

# Transform Taxa counts to relative abundance
physeq_filtered_Family_relabun <- transform_sample_counts(physeq_filtered_Family , function(OTU) OTU/sum(OTU) * 100)

#Then extract the data from the phyloseq object:
taxa_physeq_filtered_Family_relabun <- psmelt(physeq_filtered_Family_relabun)

# Sample_type

pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_5PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 5) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_0dot5_PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 0.5) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()

pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_0dot05_PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 0.05) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()



pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_0dot005_PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 0.005) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_0dot0005_PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 0.0005) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()





pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_0dot00005_PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type, y = Abundance, fill = Sample_type)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 0.00005) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


# Sample_type_DETAILED



pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type_DETAILED, y = Abundance, fill = Sample_type_DETAILED)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_5PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type_DETAILED, y = Abundance, fill = Sample_type_DETAILED)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 5) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_0dot5_PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type_DETAILED, y = Abundance, fill = Sample_type_DETAILED)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 0.5) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()

pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_0dot05_PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type_DETAILED, y = Abundance, fill = Sample_type_DETAILED)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 0.05) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()



pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_0dot005_PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type_DETAILED, y = Abundance, fill = Sample_type_DETAILED)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 0.005) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_0dot0005_PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type_DETAILED, y = Abundance, fill = Sample_type_DETAILED)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 0.0005) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()





pdf("V3_PLOT_ACCORDING_TO_COND_BOXPLOT_TREATMENT_Family_less_than_0dot00005_PERC.pdf", width=140, height = 60)

taxa_physeq_filtered_Family_relabun %>% 
  ggplot(aes(x =Sample_type_DETAILED, y = Abundance, fill = Sample_type_DETAILED)) +
  geom_boxplot() +
  labs(x = "",
       y = "Relative Abundance",
       title = "Family Relative Abundance") +
  ylim(0, 0.00005) +
  facet_wrap(TREATMENT~Family, scales= "free_x", nrow=4) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

dev.off()


######################
######################


###################################
#TEST PLOT ONLY SPECIFIC Phylum
###################################

physeq_filtered.rel = transform_sample_counts(physeq_filtered, function(x) x/sum(x)*100)

some_Phylum <- physeq_filtered.rel
some_Phylum <- tax_glom(some_Phylum, "Phylum")
#tax_table(some_Phylum)[,6]

some_Phylum <- subset_taxa(some_Phylum, Phylum == "Bactroidota"| Phylum == "Firmicutes" 
                        | Phylum == "Proteobacteria" | Phylum == "Actinobacteriota" | Phylum == "Fusobacteriota" | Phylum == "Verrucomicrobiota")

some_Phylum <- psmelt(some_Phylum)

#Sample_type

# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_NO_SCALE_Sample_type.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  #ylim(0, 5) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()


# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_10_ABUND_Sample_type.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 10) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_5_ABUND_Sample_type.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 5) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_1_ABUND_Sample_type.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 1) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_01_ABUND_Sample_type.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 0.01) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()


###################################
#END OF TEST PLOT ONLY SPECIFIC Phylum
###################################




###################################
#TEST PLOT ONLY SPECIFIC Family
###################################


some_Family <- physeq_filtered.rel
some_Family <- tax_glom(some_Family, "Family")
#tax_table(some_Family)[,6]

some_Family <- subset_taxa(some_Family, Family == "Enterobacteriaceae"| Family == "Oscillospiraceae" 
                        |Family == "Lachnospiraceae" | Family == "Coriobacteriaceae"| Family == "Fusobacteriaceae" |
                          Family == "Streptococcaceae" | Family == "Sutterellaceae" | Family == "Ethanoligenenaceae" 
                         | Family == "Morganellaceae" | Family == "Erysipelatoclostridiaceae" | Family == "Enterococcaceae"
                         )
some_Family <- psmelt(some_Family)

# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_NO_SCALE_Sample_type.pdf", width = 30, height = 40)
ggplot(some_Family, aes(x = CONCENTRATION, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type), shape=16,size = 2, position=position_jitter(0.2)) +
  #ylim(0, 5) +
  facet_wrap(TREATMENT~Family, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()


# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_MAX_10_ABUND_Sample_type.pdf", width = 30, height = 40)
ggplot(some_Family, aes(x = CONCENTRATION, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 10) +
  facet_wrap(TREATMENT~Family, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_MAX_5_ABUND_Sample_type.pdf", width = 30, height = 40)
ggplot(some_Family, aes(x = CONCENTRATION, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 5) +
  facet_wrap(TREATMENT~Family, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_MAX_1_ABUND_Sample_type.pdf", width = 30, height = 40)
ggplot(some_Family, aes(x = CONCENTRATION, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 1) +
  facet_wrap(TREATMENT~Family, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()



###################################
#END OF TEST PLOT ONLY SPECIFIC Family
###################################




###################################
#TEST PLOT ONLY SPECIFIC GENERA
###################################

physeq_filtered.rel = transform_sample_counts(physeq_filtered_FECES, function(x) x/sum(x)*100)

some_Genus <- physeq_filtered.rel
some_Genus <- tax_glom(some_Genus, "Genus")
#tax_table(some_Genus)[,6]

some_Genus <- subset_taxa(some_Genus, Genus == "Prevotella"| Genus == "Veillonella" 
                        |Genus == "Blautia" |
                          Genus == "Streptococcus"| Genus == "Faecalibacterium" |
                          Genus == "Roseburia" | Genus == "Coprococcus" |
                          Genus == "Veillonella" | Genus == "Butyricicoccus" |
                          Genus == "Eubacterium" | Genus == "[Eubacterium] hallii group" | 
                         Genus == "Butyrivibrio" | Genus == "Lachnospira" |
                          Genus == "Haemophilus" | Genus == "Akkermansia" |
                         Genus == "Erysipelatoclostridium" | Genus == "Roseburia" | Genus== "Elusimicrobium"
                         )
some_Genus <- psmelt(some_Genus)

# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_NO_SCALE_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Genus, aes(x = TREATMENT, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = TREATMENT), shape=16,size = 2, position=position_jitter(0.2)) +
  #ylim(0, 5) +
  facet_wrap(~Genus, ncol = 6) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()


# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_MAX_10_ABUND_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Genus, aes(x = TREATMENT, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = TREATMENT), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 10) +
  facet_wrap(~Genus, ncol = 6) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_MAX_5_ABUND_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Genus, aes(x = TREATMENT, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = TREATMENT), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 5) +
  facet_wrap(~Genus, ncol = 6) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_MAX_1_ABUND_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Genus, aes(x = TREATMENT, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = TREATMENT), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 1) +
  facet_wrap(~Genus, ncol = 6) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()



###################################
#END OF TEST PLOT ONLY SPECIFIC Genus
###################################

# NOW SAME FOR TREATMENT_DETAILED



#Sample_type_DETAILED


#some_Phylum <- physeq_filtered.rel
#some_Phylum <- tax_glom(some_Phylum, "Phylum")
#tax_table(some_Phylum)[,6]

#some_Phylum <- subset_taxa(some_Phylum, Phylum == "Bactroidota"| Phylum == "Firmicutes" 
#                        | Phylum == "Proteobacteria" | Phylum == "Actinobacteriota" | Phylum == "Fusobacteriota" | Phylum == #"Verrucomicrobiota")

#some_Phylum <- psmelt(some_Phylum)

# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify

pdf("Selected_Phylum_NO_SCALE_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = TREATMENT), shape=16,size = 2, position=position_jitter(0.2)) +
  #ylim(0, 5) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()


# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_10_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = TREATMENT), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 10) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_5_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = TREATMENT), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 5) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_1_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = TREATMENT), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 1) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_01_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = TREATMENT), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 0.01) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()


###################################
#END OF TEST PLOT ONLY SPECIFIC Phylum
###################################




###################################
#TEST PLOT ONLY SPECIFIC Family
###################################


#some_Family <- physeq_filtered.rel
#some_Family <- tax_glom(some_Family, "Family")
#tax_table(some_Family)[,6]

#some_Family <- subset_taxa(some_Family, Family == "Enterobacteriaceae"| Family == "Oscillospiraceae" 
#                        |Family == "Lachnospiraceae" | Family == "Coriobacteriaceae"| Family == "Fusobacteriaceae" |
#                          Family == "Streptococcaceae" | Family == "Sutterellaceae" | Family == "Ethanoligenenaceae" 
#                         | Family == "Morganellaceae" | Family == "Erysipelatoclostridiaceae" | Family == "Enterococcaceae"
#                         )
#some_Family <- psmelt(some_Family)

# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_NO_SCALE_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Family, aes(x = CONCENTRATION, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type_DETAILED), shape=16,size = 2, position=position_jitter(0.2)) +
  #ylim(0, 5) +
  facet_wrap(TREATMENT~Family, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()


# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_MAX_10_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Family, aes(x = CONCENTRATION, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type_DETAILED), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 10) +
  facet_wrap(TREATMENT~Family, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_MAX_5_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Family, aes(x = CONCENTRATION, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type_DETAILED), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 5) +
  facet_wrap(TREATMENT~Family, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_MAX_1_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Family, aes(x = CONCENTRATION, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type_DETAILED), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 1) +
  facet_wrap(TREATMENT~Family, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()



###################################
#END OF TEST PLOT ONLY SPECIFIC Family
###################################





###################################
#TEST PLOT ONLY SPECIFIC GENERA
###################################

#physeq_filtered.rel = transform_sample_counts(physeq_filtered, function(x) x/sum(x)*100)

#some_Genus <- physeq_filtered.rel
#some_Genus <- tax_glom(some_Genus, "Genus")
#tax_table(some_Genus)[,6]

#some_Genus <- subset_taxa(some_Genus, Genus == "Prevotella"| Genus == "Veillonella" 
#                        |Genus == "Blautia" |
#                          Genus == "Streptococcus"| Genus == "Faecalibacterium" |
#                          Genus == "Roseburia" | Genus == "Coprococcus" |
#                          Genus == "Veillonella" | Genus == "Butyricicoccus" |
#                          Genus == "Eubacterium" | Genus == "[Eubacterium] hallii group" | 
#                         Genus == "Butyrivibrio" | Genus == "Lachnospira" |
#                          Genus == "Haemophilus" | Genus == "Akkermansia" |
#                         Genus == "Erysipelatoclostridium" | Genus == "Roseburia" | Genus== "Elusimicrobium"
#                         )
#some_Genus <- psmelt(some_Genus)

# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_NO_SCALE_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Genus, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  #ylim(0, 5) +
  facet_wrap(~Genus, ncol = 5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()


# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_MAX_10_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Genus, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 10) +
  facet_wrap(~Genus, ncol = 5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_MAX_5_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Genus, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 5) +
  facet_wrap(~Genus, ncol = 5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_MAX_1_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Genus, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 1) +
  facet_wrap(~Genus, ncol = 5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()



###################################
#END OF TEST PLOT ONLY SPECIFIC Genus
###################################





#Sample_type_DETAILED FACET TREATMENT


#some_Phylum <- physeq_filtered.rel
#some_Phylum <- tax_glom(some_Phylum, "Phylum")
#tax_table(some_Phylum)[,6]

#some_Phylum <- subset_taxa(some_Phylum, Phylum == "Bactroidota"| Phylum == "Firmicutes" 
#                        | Phylum == "Proteobacteria" | Phylum == "Actinobacteriota" | Phylum == "Fusobacteriota" | Phylum == #"Verrucomicrobiota")

#some_Phylum <- psmelt(some_Phylum)

# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify

pdf("Selected_Phylum_NO_SCALE_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  #ylim(0, 5) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()


# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_10_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 10) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_5_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 5) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_1_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 1) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Phylum_MAX_01_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 30, height = 40)
ggplot(some_Phylum, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 0.01) +
  facet_wrap(TREATMENT~Phylum, ncol =5) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Phylum") +
  ylab("Relative abundance")
dev.off()


###################################
#END OF TEST PLOT ONLY SPECIFIC Phylum
###################################




###################################
#TEST PLOT ONLY SPECIFIC Family
###################################


#some_Family <- physeq_filtered.rel
#some_Family <- tax_glom(some_Family, "Family")
#tax_table(some_Family)[,6]

#some_Family <- subset_taxa(some_Family, Family == "Enterobacteriaceae"| Family == "Oscillospiraceae" 
                     #   |Family == "Lachnospiraceae" | Family == "Coriobacteriaceae"| Family == "Fusobacteriaceae" |
                      #    Family == "Streptococcaceae" | Family == "Sutterellaceae" | Family == "Ethanoligenenaceae" 
                       #  | Family == "Morganellaceae" | Family == "Erysipelatoclostridiaceae" | Family == "Enterococcaceae"
                        # )
#some_Family <- psmelt(some_Family)

# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_NO_SCALE_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 40, height = 40)
ggplot(some_Family, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type_DETAILED), shape=16,size = 2, position=position_jitter(0.2)) +
  #ylim(0, 5) +
  facet_wrap(TREATMENT~Family, ncol =11) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()


# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_MAX_10_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 40, height = 40)
ggplot(some_Family, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type_DETAILED), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 10) +
  facet_wrap(TREATMENT~Family, ncol =11) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_MAX_5_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 40, height = 40)
ggplot(some_Family, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type_DETAILED), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 5) +
  facet_wrap(TREATMENT~Family, ncol =11) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Family_MAX_1_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 40, height = 40)
ggplot(some_Family, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = Sample_type_DETAILED), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 1) +
  facet_wrap(TREATMENT~Family, ncol =11) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Family") +
  ylab("Relative abundance")
dev.off()



###################################
#END OF TEST PLOT ONLY SPECIFIC Family
###################################





###################################
#TEST PLOT ONLY SPECIFIC GENERA
###################################

#physeq_filtered.rel = transform_sample_counts(physeq_filtered, function(x) x/sum(x)*100)

#some_Genus <- physeq_filtered.rel
#some_Genus <- tax_glom(some_Genus, "Genus")
#tax_table(some_Genus)[,6]

#some_Genus <- subset_taxa(some_Genus, Genus == "Prevotella"| Genus == "Veillonella" 
                     #   |Genus == "Blautia" |
                    #      Genus == "Streptococcus"| Genus == "Faecalibacterium" |
                    #      Genus == "Roseburia" | Genus == "Coprococcus" |
                    #      Genus == "Veillonella" | Genus == "Butyricicoccus" |
                    #      Genus == "Eubacterium" | Genus == "[Eubacterium] hallii group" | 
                    #     Genus == "Butyrivibrio" | Genus == "Lachnospira" |
                    #      Genus == "Haemophilus" | Genus == "Akkermansia" |
                    #     Genus == "Erysipelatoclostridium" | Genus == "Roseburia" | Genus== "Elusimicrobium"
                    #     )
#some_Genus <- psmelt(some_Genus)

# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_NO_SCALE_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 60, height = 40)
ggplot(some_Genus, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  #ylim(0, 5) +
  facet_wrap(TREATMENT~Genus, ncol = 12) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()


# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_MAX_10_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 60, height = 40)
ggplot(some_Genus, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 10) +
  facet_wrap(TREATMENT~Genus, ncol = 12) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_MAX_5_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 60, height = 40)
ggplot(some_Genus, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 5) +
  facet_wrap(TREATMENT~Genus, ncol = 12) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()



# Plot  
# x = "", color ="", , facet~wrap and comparison list can be modify
pdf("Selected_Genus_MAX_1_ABUND_Sample_type_DETAILED_FACET_TREATMENT.pdf", width = 60, height = 40)
ggplot(some_Genus, aes(x = Sample_type_DETAILED, y = Abundance)) +
  geom_boxplot() +
  #geom_point(aes(color = TREATMENT)) +
  geom_jitter(aes(colour = CONCENTRATION), shape=16,size = 2, position=position_jitter(0.2)) +
  ylim(0, 1) +
  facet_wrap(TREATMENT~Genus, ncol = 12) +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) +
  ggtitle("Relative abundance of selected Genus") +
  ylab("Relative abundance")
dev.off()



###################################
#END OF TEST PLOT ONLY SPECIFIC Genus Sample_type_DETAILED FACET TREATMENT
###################################












###################################
#MAASLIN2 TEST
###################################

df_input_metadata = read.table(file             = "Metadata.txt", 
                               header           = TRUE, 
                               sep              = "\t", 
                               row.names        = 1,
                               stringsAsFactors = FALSE)
#View(df_input_metadata)




#Preparing the data


# Phylum Level

physeq_filtered_rel_abund_Phylum <- physeq_filtered %>% aggregate_taxa(level = "Phylum") %>% microbiome::transform(transform = "compositional")

otu_table_physeq_filtered_rel_abund_Phylum  <- otu_table(physeq_filtered_rel_abund_Phylum)

# Order Level


physeq_filtered_rel_abund_Order <- physeq_filtered %>% aggregate_taxa(level = "Order") %>% microbiome::transform(transform = "compositional")

otu_table_physeq_filtered_rel_abund_Order  <- otu_table(physeq_filtered_rel_abund_Order)

# CLass level


physeq_filtered_rel_abund_Class <- physeq_filtered %>% aggregate_taxa(level = "Class") %>% microbiome::transform(transform = "compositional")

otu_table_physeq_filtered_rel_abund_Class  <- otu_table(physeq_filtered_rel_abund_Class)

#Family level 

physeq_filtered_rel_abund_Family <- physeq_filtered %>% aggregate_taxa(level = "Family") %>% microbiome::transform(transform = "compositional")

otu_table_physeq_filtered_rel_abund_Family  <- otu_table(physeq_filtered_rel_abund_Family)

# Genus level

physeq_filtered_rel_abund_Genus <- physeq_filtered %>% aggregate_taxa(level = "Genus") %>% microbiome::transform(transform = "compositional")

otu_table_physeq_filtered_rel_abund_Genus  <- otu_table(physeq_filtered_rel_abund_Genus)


# Species level

physeq_filtered_rel_abund_Species <- physeq_filtered %>% aggregate_taxa(level = "Species") %>% microbiome::transform(transform = "compositional")

otu_table_physeq_filtered_rel_abund_Species  <- otu_table(physeq_filtered_rel_abund_Species)



#Phylum level
fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Phylum, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_Phylum", 
                     fixed_effects  = c("Sample_type"),
                     reference      = c("Sample_type,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)

# Order level

fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Order, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_Order", 
                     fixed_effects  = c("Sample_type"),
                     reference      = c("Sample_type,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)

# Class Level


fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Class, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_Class", 
                     fixed_effects  = c("Sample_type"),
                     reference      = c("Sample_type,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)


# Family Level



fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Family, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_Family", 
                     fixed_effects  = c("Sample_type"),
                     reference      = c("Sample_type,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)

# Genus Level


fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Genus, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_Genus", 
                     fixed_effects  = c("Sample_type"),
                     reference      = c("Sample_type,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)


# Sample_type Species Level

fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Species, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_Species", 
                     fixed_effects  = c("Sample_type"),
                     reference      = c("Sample_type,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)


#BILE_STATES ANALYSIS

#Phylum level
fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Phylum, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_BILE_STATES_Phylum", 
                     fixed_effects  = c("BILE_STATES"),
                     reference      = c("BILE_STATES,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)



# Order Level


fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Order, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_BILE_STATES_Order", 
                     fixed_effects  = c("BILE_STATES"),
                     reference      = c("BILE_STATES,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)

# Class Level


fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Class, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_BILE_STATES_Class", 
                     fixed_effects  = c("BILE_STATES"),
                     reference      = c("BILE_STATES,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)


# Family Level


fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Family, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_BILE_STATES_Family", 
                     fixed_effects  = c("BILE_STATES"),
                     reference      = c("BILE_STATES,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)

# Genus Level


fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Genus, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_BILE_STATES_Genus", 
                     fixed_effects  = c("BILE_STATES"),
                     reference      = c("BILE_STATES,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)


# BILE_STATES Species Level


fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Species, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_BILE_STATES_Species", 
                     fixed_effects  = c("BILE_STATES"),
                     reference      = c("BILE_STATES,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)

# Sample_type_DETAILED


#Phylum level
fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Phylum, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_DETAILED_Phylum", 
                     fixed_effects  = c("Sample_type_DETAILED"),
                     reference      = c("Sample_type_DETAILED,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)

# Order level

fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Order, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_DETAILED_Order", 
                     fixed_effects  = c("Sample_type_DETAILED"),
                     reference      = c("Sample_type_DETAILED,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)

# Class Level


fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Class, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_DETAILED_Class", 
                     fixed_effects  = c("Sample_type_DETAILED"),
                     reference      = c("Sample_type_DETAILED,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)


# Family Level



fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Family, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_DETAILED_Family", 
                     fixed_effects  = c("Sample_type_DETAILED"),
                     reference      = c("Sample_type_DETAILED,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)

# Genus Level


fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Genus, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_DETAILED_Genus", 
                     fixed_effects  = c("Sample_type_DETAILED"),
                     reference      = c("Sample_type_DETAILED,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)


# Sample_type_DETAILED Species Level

fit_data = Maaslin2(input_data     = otu_table_physeq_filtered_rel_abund_Species, 
                     input_metadata = df_input_metadata, 
                     min_prevalence = 0.1,
                     min_abundance  = 0.0001,
                     normalization  = "NONE",
                     output         = "Maaslin2_Analysis_Sample_type_DETAILED_Species", 
                     fixed_effects  = c("Sample_type_DETAILED"),
                     reference      = c("Sample_type_DETAILED,CTRL"),
                     random_effects = c("TREATMENT"),
                     cores=10)




###################################
# END OF MAASLIN2 TEST
###################################



##################################################

#HEATMAP OF ABUNDANCE (BUT ONLY PRESENCE/ABSENCE)!!!


# make a baTREATMENT heatmap
#pdf("BATREATMENT_PRESENCE_ABSENCE_HEATMAP_ALL_SAMPLES.pdf", width = 20, height=20)
pdf("BATREATMENT_HEATMAP_ALL_SAMPLES.pdf")

phyloseq::plot_heatmap(physeq_filtered, "PCoA", "bray", "TREATMENT", low="#66CCFF", high="#000033", na.value="white", title = "PCoA-Bray")
dev.off()



#I WILL TRY TO DO A NORMAL HEAT MAP, BUT COLOR TOTAL BLACK WHEN HIGHER THAN ZERO!! GENUS


# Relative abundance for each group
vb_1 <- subset_samples(physeq_filtered)

# Percentage
vb_1 <- transform_sample_counts(vb_1, function(x) 100*x/sum(x))

# Agglomerate as a taxonomic rank
vb_1 <- tax_glom(vb_1, "Genus")

# Prune top X genera (only show top x values)
vb_1 <- prune_taxa(names(sort(taxa_sums(vb_1), TRUE)[1:100]), vb_1)

# Merging subest phyloseq - Genus level agglomeration
physeq_vb <- tax_glom(vb_1, "Genus")

pdf("physeq_filtered_heat_final_REL_ABUND_PHYLUM_TEST_2_Genus.pdf", width = 30, height = 10)

phyloseq::plot_heatmap(physeq_vb, "NMDS", "jaccard", "Sample_type", low="#66CCFF", high="#000033", na.value="white", title = "NMDS-jaccard", "Genus")
dev.off()

pdf("physeq_filtered_heat_final_REL_ABUND_PHYLUM_TEST_3_Genus.pdf", width = 30, height = 10)

phyloseq::plot_heatmap(physeq_vb, "NMDS", "jaccard", "TREATMENT", low="#66CCFF", high="#000033", na.value="white", title = "NMDS-jaccard", "Genus")
dev.off()





#I WILL TRY TO DO A NORMAL HEAT MAP, BUT COLOR TOTAL BLACK WHEN HIGHER THAN ZERO!! CLASS


# Relative abundance for each group
vb_1 <- subset_samples(physeq_filtered)

# Percentage
vb_1 <- transform_sample_counts(vb_1, function(x) 100*x/sum(x))

# Agglomerate as a taxonomic rank
vb_1 <- tax_glom(vb_1, "Class")

# Prune top X genera (only show top x values)
vb_1 <- prune_taxa(names(sort(taxa_sums(vb_1), TRUE)[1:100]), vb_1)

# Merging subest phyloseq - Class level agglomeration
physeq_vb <- tax_glom(vb_1, "Class")

pdf("physeq_filtered_heat_final_REL_ABUND_PHYLUM_TEST_2_Class.pdf", width = 30, height = 10)

phyloseq::plot_heatmap(physeq_vb, "NMDS", "jaccard", "Sample_type", low="#66CCFF", high="#000033", na.value="white", title = "NMDS-jaccard", "Class")
dev.off()

pdf("physeq_filtered_heat_final_REL_ABUND_PHYLUM_TEST_3_Class.pdf", width = 30, height = 10)

phyloseq::plot_heatmap(physeq_vb, "NMDS", "jaccard", "TREATMENT", low="#66CCFF", high="#000033", na.value="white", title = "NMDS-jaccard", "Class")
dev.off()



#I WILL TRY TO DO A NORMAL HEAT MAP, BUT COLOR TOTAL BLACK WHEN HIGHER THAN ZERO!! PHYLUM


# Relative abundance for each group
vb_1 <- subset_samples(physeq_filtered)

# Percentage
vb_1 <- transform_sample_counts(vb_1, function(x) 100*x/sum(x))

# Agglomerate as a taxonomic rank
vb_1 <- tax_glom(vb_1, "Phylum")

# Prune top X genera (only show top x values)
vb_1 <- prune_taxa(names(sort(taxa_sums(vb_1), TRUE)[1:100]), vb_1)

# Merging subest phyloseq - Phylum level agglomeration
physeq_vb <- tax_glom(vb_1, "Phylum")

pdf("physeq_filtered_heat_final_REL_ABUND_PHYLUM_TEST_2_Phylum.pdf", width = 30, height = 10)

phyloseq::plot_heatmap(physeq_vb, "NMDS", "jaccard", "TREATMENT", low="#66CCFF", high="#000033", na.value="white", title = "NMDS-jaccard", "Phylum") +
theme(axis.text.x = element_text(angle = -80, hjust = 0, vjust=0.5)) 
dev.off()







#################################################
#################################################
#################################################



#################################################




##################### TEST JUL RELATIVE ABUND working
###### PHYLUM

physeq_filtered.rel = transform_sample_counts(physeq_filtered, function(OTU) (OTU/sum(OTU))*100)


get_taxa(physeq_filtered.rel, sample_names(physeq)[100])


physeq_filtered.rel = transform_sample_counts(physeq_filtered, function(OTU) (OTU/sum(OTU))*100)
# agglomerate taxa
glom <- tax_glom(physeq_filtered.rel, taxrank = 'Phylum', NArm = FALSE)


physeq_filtered.melt <- psmelt(glom)


physeq_filtered.melt <- physeq_filtered.melt %>%
  group_by(TREATMENT,Sample_type,Phylum) %>%
  mutate(median=median(Abundance))
# select group median > 1
keep <- unique(physeq_filtered.melt$Phylum[physeq_filtered.melt$median > 1])
physeq_filtered.melt$Phylum[!(physeq_filtered.melt$Phylum %in% keep)] <- "< 1%"
#to get the same rows together
physeq_filtered.melt_sum <- physeq_filtered.melt %>%
  group_by(Sample_type, TREATMENT, CONCENTRATION, Phylum) %>%
  summarise(Abundance=sum(Abundance)/3)


# change to character for easy-adjusted level
physeq_filtered.melt_sum$Phylum <- as.character(physeq_filtered.melt_sum$Phylum)



#View(physeq_filtered.melt.test)

pdf("BarPlot_RelativeAbundance_TREATMENTCOMs_Phylum_Sample_type_V2.pdf",  width=20, height=60)

ggplot(physeq_filtered.melt_sum, aes(x = CONCENTRATION, y = Abundance)) + 
  geom_bar(stat = "identity", aes(fill=Phylum)) + 
  labs(x="", y="%") +
  facet_wrap(TREATMENT ~ Sample_type, scales= "free_x", ncol=5) +
 theme_classic() + 
  theme(strip.background = element_blank(), 
        axis.text.x.bottom = element_text(angle = -90))
dev.off()












######CLASS

physeq_filtered.rel = transform_sample_counts(physeq_filtered, function(x) x/sum(x)*100)
# agglomerate taxa
glom <- tax_glom(physeq_filtered.rel, taxrank = 'Class', NArm = FALSE)
physeq_filtered.melt <- psmelt(glom)
# change to character for easy-adjusted level
physeq_filtered.melt$Class <- as.character(physeq_filtered.melt$Class)

physeq_filtered.melt <- physeq_filtered.melt %>%
  group_by(Sample_type, Class) %>%
  mutate(median=median(Abundance))
# select group median > 1
keep <- unique(physeq_filtered.melt$Class[physeq_filtered.melt$median > 1])
physeq_filtered.melt$Class[!(physeq_filtered.melt$Class %in% keep)] <- "< 1%"
#to get the same rows together
physeq_filtered.melt_sum <- physeq_filtered.melt %>%
  group_by(TREATMENT,Sample_type,Class) %>%
  summarise(Abundance=sum(Abundance))

pdf("BarPlot_RelativeAbundance_TREATMENTCOMs_Class_Sample_type_V2.pdf",  width=20, height=10)

ggplot(physeq_filtered.melt_sum, aes(x = TREATMENT, y = Abundance, fill = Class)) + 
  geom_bar(stat = "identity", aes(fill=Class)) + 
  labs(x="", y="%") +
  facet_wrap(TREATMENT~Sample_type, scales= "free_x", nrow=1) +
 theme_classic() + 
  theme(strip.background = element_blank(), 
        axis.text.x.bottom = element_text(angle = -90))
dev.off()



######FAMILY

pdf("BarPlot_RelativeAbundance_TREATMENTCOMs_Family_Sample_type_V2.pdf",  width=20, height=10)
physeq_filtered.rel = transform_sample_counts(physeq_filtered, function(x) x/sum(x)*100)
# agglomerate taxa
glom <- tax_glom(physeq_filtered.rel, taxrank = 'Family', NArm = FALSE)
physeq_filtered.melt <- psmelt(glom)
# change to character for easy-adjusted level
physeq_filtered.melt$Family <- as.character(physeq_filtered.melt$Family)

physeq_filtered.melt <- physeq_filtered.melt %>%
  group_by(Sample_type, Family) %>%
  mutate(median=median(Abundance))
# select group median > 1
keep <- unique(physeq_filtered.melt$Family[physeq_filtered.melt$median > 1])
physeq_filtered.melt$Family[!(physeq_filtered.melt$Family %in% keep)] <- "< 1%"
#to get the same rows together
physeq_filtered.melt_sum <- physeq_filtered.melt %>%
  group_by(TREATMENT,Sample_type,Family) %>%
  summarise(Abundance=sum(Abundance))

ggplot(physeq_filtered.melt_sum, aes(x = TREATMENT, y = Abundance, fill = Family)) + 
  geom_bar(stat = "identity", aes(fill=Family)) + 
  labs(x="", y="%") +
  facet_wrap(TREATMENT~Sample_type, scales= "free_x", nrow=1) +
 theme_classic() + 
  theme(strip.background = element_blank(), 
        axis.text.x.bottom = element_text(angle = -90))
dev.off()



######GENUS

pdf("BarPlot_RelativeAbundance_TREATMENTCOMs_Genus_Sample_type_V2.pdf",  width=20, height=10)
physeq_filtered.rel = transform_sample_counts(physeq_filtered, function(x) x/sum(x)*100)
# agglomerate taxa
glom <- tax_glom(physeq_filtered.rel, taxrank = 'Genus', NArm = FALSE)
physeq_filtered.melt <- psmelt(glom)
# change to character for easy-adjusted level
physeq_filtered.melt$Genus <- as.character(physeq_filtered.melt$Genus)

physeq_filtered.melt <- physeq_filtered.melt %>%
  group_by(Sample_type, Genus) %>%
  mutate(median=median(Abundance))
# select group median > 1
keep <- unique(physeq_filtered.melt$Genus[physeq_filtered.melt$median > 1])
physeq_filtered.melt$Genus[!(physeq_filtered.melt$Genus %in% keep)] <- "< 1%"
#to get the same rows together
physeq_filtered.melt_sum <- physeq_filtered.melt %>%
  group_by(TREATMENT,Sample_type,Genus) %>%
  summarise(Abundance=sum(Abundance))

ggplot(physeq_filtered.melt_sum, aes(x = TREATMENT, y = Abundance, fill = Genus)) + 
  geom_bar(stat = "identity", aes(fill=Genus)) + 
  labs(x="", y="%") +
  facet_wrap(TREATMENT~Sample_type, scales= "free_x", nrow=1) +
 theme_classic() + 
  theme(strip.background = element_blank(), 
        axis.text.x.bottom = element_text(angle = -90))
dev.off()



######SPECIES

pdf("BarPlot_RelativeAbundance_TREATMENTCOMs_Species_Sample_type_V2.pdf",  width=20, height=10)
physeq_filtered.rel = transform_sample_counts(physeq_filtered, function(x) x/sum(x)*100)
# agglomerate taxa
glom <- tax_glom(physeq_filtered.rel, taxrank = 'Species', NArm = FALSE)
physeq_filtered.melt <- psmelt(glom)
# change to character for easy-adjusted level
physeq_filtered.melt$Species <- as.character(physeq_filtered.melt$Species)

physeq_filtered.melt <- physeq_filtered.melt %>%
  group_by(Sample_type, Species) %>%
  mutate(median=median(Abundance))
# select group median > 1
keep <- unique(physeq_filtered.melt$Species[physeq_filtered.melt$median > 1])
physeq_filtered.melt$Species[!(physeq_filtered.melt$Species %in% keep)] <- "< 1%"
#to get the same rows together
physeq_filtered.melt_sum <- physeq_filtered.melt %>%
  group_by(TREATMENT,Sample_type,Species) %>%
  summarise(Abundance=sum(Abundance))

ggplot(physeq_filtered.melt_sum, aes(x = TREATMENT, y = Abundance, fill = Species)) + 
  geom_bar(stat = "identity", aes(fill=Species)) + 
  labs(x="", y="%") +
  facet_wrap(TREATMENT~Sample_type, scales= "free_x", nrow=1) +
 theme_classic() + 
  theme(strip.background = element_blank(), 
        axis.text.x.bottom = element_text(angle = -90))
dev.off()






#######################



#Make a genus vs log2FC plot of the significant OTUs:

#Adding pseudo-count of one to avoid problems with zero in Deseq analysis

# Create a copy of the physeq_filtered that we want to keep intact

physeq_filtered_copy <- physeq_filtered

View(physeq_filtered_copy)

otu_table(physeq_filtered_copy) <- otu_table(physeq_filtered_copy) + 1

#View(otu_table(physeq_filtered_copy))

sample_data(physeq_filtered_copy)$Sample_type <- as.factor(sample_data(physeq_filtered_copy)$Sample_type)

#Convert the phyloseq object to a DESeqDataSet and run DESeq2:

ds = phyloseq_to_deseq2(physeq_filtered_copy, ~ Sample_type)

#adding a pseudcount for aving no zero

# The next step takes a bit of time
ds = DESeq(ds)
View(ds)
#Extract the result table from the ds object usind the DESeq2 function results and filter the OTUs using a False Discovery Rate (FDR) cutoff of 0.01. In this example we return the significantly differentially abundant OTU between the seasons “Spring” and “Fall”:


rm(res_sig)
alpha = 0.05
res = results(ds, contrast=c("Sample_type", "Lithocholic acid", "CTRL"), alpha=alpha)
res = res[order(res$padj, na.last=NA), ]
res_sig = res[(res$padj < alpha), ]
res_sig


res_sig = cbind(as(res_sig, "data.frame"), as(tax_table(physeq_filtered_copy)[rownames(res_sig), ], "matrix"))


View(res_sig)


#### To order the graph according to log2foldchange Genus level

x = tapply(res_sig$log2FoldChange, res_sig$Genus, function(x) max(x))
x = sort(x, TRUE)
res_sig$Genus = factor(as.character(res_sig$Genus), levels=names(x))


pdf("Genus_vs_log2FC_significant_OTUs_TREATMENT_vs_FECES.pdf",  width=20, height=10)
ggplot(res_sig, aes(x=Genus, y=log2FoldChange, color=Class)) +
    geom_jitter(size=3, width = 0.2) +
    theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))
dev.off()



#######################

######################

# DESeq2 Differential abundance analysis with heatmap (https://www.yanh.org/2021/01/01/microbiome-r/)
#### GENUS level

sample_data(physeq_filtered_copy)$Sample_type <- as.factor(sample_data(physeq_filtered_copy)$Sample_type) # factorize for DESeq2

physeq_filtered_copy.taxa <- tax_glom(physeq_filtered_copy, taxrank = 'Genus', NArm = TRUE)


# pairwise comparison between BILE_STATES and CONTROL
physeq_filtered_copy.taxa.sub <- subset_samples(physeq_filtered_copy.taxa, Sample_type %in% c("Lithocholic acid", "CTRL"))
# filter sparse features, with > 90% zeros
physeq_filtered_copy.taxa.pse.sub <- prune_taxa(rowSums(otu_table(physeq_filtered_copy.taxa.sub) == 0) < ncol(otu_table(physeq_filtered_copy.taxa.sub)) * 0.9, physeq_filtered_copy.taxa.sub)
physeq_filtered_copy_ds = phyloseq_to_deseq2(physeq_filtered_copy.taxa.pse.sub, ~ Sample_type)
# use alternative estimator on a TREATMENT of "every gene contains a sample with a zero"
ds <- estimateSizeFactors(physeq_filtered_copy_ds, type="poscounts")
ds = DESeq(ds, test="Wald", fitType="parametric")
alpha = 0.05 
res = results(ds, alpha=alpha)
res = res[order(res$padj, na.last=NA), ]
taxa_sig = rownames(res[1:40, ]) # select bottom 40 with lowest p.adj values
physeq_filtered_copy.taxa.rel <- transform_sample_counts(physeq_filtered_copy, function(x) x/sum(x)*100)
physeq_filtered_copy.taxa.rel.sig <- prune_taxa(taxa_sig, physeq_filtered_copy.taxa.rel)
# Only keep "Lithocholic acid", "CTRL" samples
physeq_filtered_copy.taxa.rel.sig <- prune_samples(colnames(otu_table(physeq_filtered_copy.taxa.pse.sub)), physeq_filtered_copy.taxa.rel.sig)



matrix <- as.matrix(data.frame(otu_table(physeq_filtered_copy.taxa.rel.sig)))
rownames(matrix) <- as.character(tax_table(physeq_filtered_copy.taxa.rel.sig)[, "Genus"])
metadata_sub <- data.frame(sample_data(physeq_filtered_copy.taxa.rel.sig))
# Define the annotation color for columns and rows
annotation_col = data.frame(
    BILE_STATES = as.factor(metadata_sub$BILE_STATES), 
    `Sample_type` = as.factor(metadata_sub$Sample_type), 
    check.names = FALSE
)
rownames(annotation_col) = rownames(metadata_sub)

annotation_row = data.frame(
    Class = as.factor(tax_table(physeq_filtered_copy.taxa.rel.sig)[, "Class"])
)
rownames(annotation_row) = rownames(matrix)

# ann_color should be named vectors
Class_col = RColorBrewer::brewer.pal(length(levels(annotation_row$Class)), "Paired")
names(Class_col) = levels(annotation_row$Class)
ann_colors = list(
    `Sample_type` = c("CTRL" = "orange", "Lithocholic acid" = "brown"),
    Class = Class_col
)


pdf("DESeq2_Differential_Abundance_analysis_Genus_CTRL_vs_Lithocholic.pdf",  width=15, height=10)
ComplexHeatmap::pheatmap(matrix, scale= "row", 
                         annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         annotation_colors = ann_colors)
dev.off()



sample_data(physeq_filtered_copy)$Sample_type <- as.factor(sample_data(physeq_filtered_copy)$Sample_type) # factorize for DESeq2

physeq_filtered_copy.taxa <- tax_glom(physeq_filtered_copy, taxrank = 'Genus', NArm = TRUE)



# pairwise comparison between BILE_STATES and CONTROL
physeq_filtered_copy.taxa.sub <- subset_samples(physeq_filtered_copy.taxa, Sample_type %in% c("Chenodeoxycholic Acid", "CTRL"))
# filter sparse features, with > 90% zeros
physeq_filtered_copy.taxa.pse.sub <- prune_taxa(rowSums(otu_table(physeq_filtered_copy.taxa.sub) == 0) < ncol(otu_table(physeq_filtered_copy.taxa.sub)) * 0.9, physeq_filtered_copy.taxa.sub)
physeq_filtered_copy_ds = phyloseq_to_deseq2(physeq_filtered_copy.taxa.pse.sub, ~ Sample_type)
# use alternative estimator on a TREATMENT of "every gene contains a sample with a zero"
ds <- estimateSizeFactors(physeq_filtered_copy_ds, type="poscounts")
ds = DESeq(ds, test="Wald", fitType="parametric")
alpha = 0.05 
res = results(ds, alpha=alpha)
res = res[order(res$padj, na.last=NA), ]
taxa_sig = rownames(res[1:40, ]) # select bottom 40 with lowest p.adj values
physeq_filtered_copy.taxa.rel <- transform_sample_counts(physeq_filtered_copy, function(x) x/sum(x)*100)
physeq_filtered_copy.taxa.rel.sig <- prune_taxa(taxa_sig, physeq_filtered_copy.taxa.rel)
# Only keep "Chenodeoxycholic acid", "CTRL" samples
physeq_filtered_copy.taxa.rel.sig <- prune_samples(colnames(otu_table(physeq_filtered_copy.taxa.pse.sub)), physeq_filtered_copy.taxa.rel.sig)



matrix <- as.matrix(data.frame(otu_table(physeq_filtered_copy.taxa.rel.sig)))
rownames(matrix) <- as.character(tax_table(physeq_filtered_copy.taxa.rel.sig)[, "Genus"])
metadata_sub <- data.frame(sample_data(physeq_filtered_copy.taxa.rel.sig))
# Define the annotation color for columns and rows
annotation_col = data.frame(
    CONCENTRATION = as.factor(metadata_sub$CONCENTRATION), 
    `Sample_type` = as.factor(metadata_sub$Sample_type), 
    check.names = FALSE
)
rownames(annotation_col) = rownames(metadata_sub)

annotation_row = data.frame(
    Class = as.factor(tax_table(physeq_filtered_copy.taxa.rel.sig)[, "Class"])
)
rownames(annotation_row) = rownames(matrix)

# ann_color should be named vectors
Class_col = RColorBrewer::brewer.pal(length(levels(annotation_row$Class)), "Paired")
names(Class_col) = levels(annotation_row$Class)
ann_colors = list(
    `Sample_type` = c("CTRL" = "orange", "Chenodeoxycholic Acid" = "brown"),
    Class = Class_col
)


pdf("DESeq2_Differential_Abundance_analysis_Genus_CTRL_vs_Chenodeoxycholic.pdf",  width=15, height=10)
ComplexHeatmap::pheatmap(matrix, scale= "row", 
                         annotation_col = annotation_col, 
                         annotation_row = annotation_row, 
                         annotation_colors = ann_colors)
dev.off()




####################################################





```{r}
# Checking for relative abundances of salivarius vs Enterococcus bacteria (boxplots)


# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)
library(ggbreak)  # For breaking the y-axis

# Import data
physeq_data <- read.table("physeq_filtered_relative_abundance.all.tsv", 
                          sep = "\t", header = TRUE, stringsAsFactors = FALSE)

# Convert ASV column to rownames if necessary
rownames(physeq_data) <- physeq_data$ASV

# Filter for Enterococcus and Salivarius, excluding NA species
filtered_data <- physeq_data %>%
  filter((Species == "salivarius" | Genus == "Enterococcus") & !is.na(Species)) %>%
  select(Species, Genus, starts_with("LYS_")) %>%
  pivot_longer(cols = starts_with("LYS_"), 
               names_to = "Condition", values_to = "Abundance") %>%
  mutate(Abundance = ifelse(Abundance == 0, 0.0001, Abundance))  # Replace 0 with pseudocount

# Get unique Enterococcus species
enterococcus_species <- filtered_data %>%
  filter(Genus == "Enterococcus") %>%
  pull(Species) %>%
  unique()

# Open PDF for output
pdf("Paired_Salivarius_Enterococcus_Boxplots.pdf", width = 8, height = 6)

# Loop through each Enterococcus species and create a separate plot
for (species in enterococcus_species) {
  
  # Subset data for the current Enterococcus species and Salivarius
  plot_data <- filtered_data %>%
    filter(Species %in% c(species, "salivarius")) %>%
    mutate(Species = factor(Species, levels = c(species, "salivarius")))  # Ensure Salivarius is always on the right

  # Create a Sample_ID column to group by matching samples
  plot_data$Sample_ID <- plot_data$Condition

  # Generate plot
  p <- ggplot(plot_data, aes(x = Species, y = Abundance, fill = Species)) +
    geom_boxplot(alpha = 0.7, outlier.shape = NA) +  # Boxplots (all samples included)
    geom_jitter(aes(group = Sample_ID), shape = 16, position = position_jitter(0.2), alpha = 0.6) +  # Data points
    geom_line(aes(group = Sample_ID, color = Sample_ID), linewidth = 0.8, alpha = 0.7) +  # Connecting lines
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    labs(title = paste("Abundance of", species, "vs Salivarius"),
         x = "Bacteria", y = "Relative Abundance (%)") +
    scale_fill_viridis_d() +  # Automatically assigns colors
    scale_color_viridis_d(guide = "none")   # Use same color scheme for lines, hide legend

  # Check if y-axis needs to be broken for better visualization
  max_y <- max(plot_data$Abundance, na.rm = TRUE)
  if (max_y > 50) {
    p <- p + scale_y_break(c(5, 50))  # Break between 5 and 50 if large differences exist
  }

  # Print the plot to the PDF
  print(p)
}

# Close PDF file
dev.off()



```





```{r}

#### USING the phyloseq object with SIC8 and SIC9 (excluding SIC10) instead of importing the tables

# Load required libraries
library(phyloseq)
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)

# Extract OTU table and convert to dataframe
otu_df <- as.data.frame(otu_table(physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES_ra))

# Extract taxonomy and metadata
tax_df <- as.data.frame(tax_table(physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES_ra))
meta_df <- as.data.frame(sample_data(physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES_ra))

# Add taxonomic information to the OTU table
otu_df$Species <- tax_df$Species  

# Pivot the OTU table to long format (keeping `fastqID` as sample identifier)
long_data <- otu_df %>%
  pivot_longer(cols = -Species, names_to = "fastqID", values_to = "Abundance")

# Merge with metadata using `fastqID`
long_data <- left_join(long_data, meta_df, by = "fastqID")

# Filter for only *Salivarius* and *Coprococcus comes*
filtered_data <- long_data %>%
  filter(Species %in% c("salivarius", "comes") & !is.na(Species)) %>%
  mutate(Abundance = ifelse(Abundance == 0, 0.0001, Abundance))  

# Ensure every `fastqID` has both species before plotting
paired_data <- filtered_data %>%
  group_by(fastqID) %>%
  filter(n_distinct(Species) == 2) %>%  
  ungroup()

# Convert from long to wide format for correct line plotting
wide_data <- paired_data %>%
  select(fastqID, Species, Abundance, TREATMENT) %>%
  pivot_wider(names_from = Species, values_from = Abundance) %>%
  drop_na()  

# Ensure correct column names
colnames(wide_data) <- c("fastqID", "TREATMENT", "Coprococcus_comes", "Salivarius")

# Open PDF for output (multiple pages)
pdf("Connected_Boxplots_Salivarius_CoprococcusComes.pdf", width = 10, height = 6)

# First Page: Full-Scale Boxplot
p1 <- ggplot(paired_data, aes(x = Species, y = Abundance, fill = Species)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +  
  geom_jitter(aes(color = TREATMENT), shape = 16, position = position_jitter(0.2), alpha = 0.6) +  
  geom_line(aes(group = fastqID, color = TREATMENT), alpha = 0.5) +
  scale_x_discrete(labels = c("Coprococcus comes", "Salivarius")) +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  labs(title = "Paired Abundance of Coprococcus comes vs Salivarius",
       x = "Bacteria", y = "Relative Abundance (%)", color = "Treatment") +
  scale_fill_viridis_d() +  
  scale_color_manual(values = c("Lysin" = "#E69F00", "NO_LYSIN" = "#56B4E9"))  

print(p1)  # Print first page

# Second Page: Zoomed-In View (Max Y = 5%)
p2 <- ggplot(paired_data, aes(x = Species, y = Abundance, fill = Species)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +  
  geom_jitter(aes(color = TREATMENT), shape = 16, position = position_jitter(0.2), alpha = 0.6) +  
  geom_line(aes(group = fastqID, color = TREATMENT), alpha = 0.5) +
  scale_x_discrete(labels = c("Coprococcus comes", "Salivarius")) +  
  ylim(0, 5) +  # Set y-axis limit to 5%
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  labs(title = "Zoomed-In: Low Abundance (Below 5%) - Coprococcus comes vs Salivarius",
       x = "Bacteria", y = "Relative Abundance (%)", color = "Treatment") +
  scale_fill_viridis_d() +  
  scale_color_manual(values = c("Lysin" = "#E69F00", "NO_LYSIN" = "#56B4E9"))  

print(p2)  # Print second page

# Close PDF file
dev.off()


```



# Load required libraries
library(phyloseq)
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)

# Extract OTU table and convert to dataframe
otu_df <- as.data.frame(otu_table(physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES_ra))

# Extract taxonomy and metadata
tax_df <- as.data.frame(tax_table(physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES_ra))
meta_df <- as.data.frame(sample_data(physeq_glom_physeq_filtered_SIC8_SIC9_NO_EB_24_48_SPECIES_ra))

# Add taxonomic information to the OTU table
otu_df$Species <- tax_df$Species  
otu_df$Genus <- tax_df$Genus  

# Pivot the OTU table to long format (keeping `fastqID` as sample identifier)
long_data <- otu_df %>%
  pivot_longer(cols = -c(Species, Genus), names_to = "fastqID", values_to = "Abundance")

# Merge with metadata using `fastqID`
long_data <- left_join(long_data, meta_df, by = "fastqID")

# Filter for *Salivarius* and all *Enterococcus* species
filtered_data <- long_data %>%
  filter(Species == "salivarius" | Genus == "Enterococcus") %>%
  mutate(Abundance = ifelse(Abundance == 0, 0.0001, Abundance))  

# Get unique Enterococcus species
enterococcus_species <- filtered_data %>%
  filter(Genus == "Enterococcus") %>%
  pull(Species) %>%
  unique()

# Open multi-page PDF for output
pdf("Connected_Boxplots_Salivarius_vs_Enterococcus.pdf", width = 10, height = 6)

# Loop through each Enterococcus species
for (species in enterococcus_species) {
  
  # Subset data for the current Enterococcus species and *Salivarius*
  paired_data <- filtered_data %>%
    filter(Species %in% c(species, "salivarius")) %>%
    group_by(fastqID) %>%
    filter(n_distinct(Species) == 2) %>%  # Ensure both species are present
    ungroup()
  
  # Convert from long to wide format for correct line plotting
  wide_data <- paired_data %>%
    select(fastqID, Species, Abundance, TREATMENT) %>%
    pivot_wider(names_from = Species, values_from = Abundance) %>%
    drop_na()  

  # Rename columns for clarity
  colnames(wide_data) <- c("fastqID", "TREATMENT", "Enterococcus", "Salivarius")

  # First Page: Full-Scale Boxplot
  p1 <- ggplot(paired_data, aes(x = Species, y = Abundance, fill = Species)) +
    geom_boxplot(alpha = 0.7, outlier.shape = NA) +  
    geom_jitter(aes(color = TREATMENT), shape = 16, position = position_jitter(0.2), alpha = 0.6) +  
    geom_line(aes(group = fastqID, color = TREATMENT), alpha = 0.5) +
    scale_x_discrete(labels = c(species, "Salivarius")) +  
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    labs(title = paste("Paired Abundance of", species, "vs Salivarius"),
         x = "Bacteria", y = "Relative Abundance (%)", color = "Treatment") +
    scale_fill_viridis_d() +  
    scale_color_manual(values = c("Lysin" = "#E69F00", "NO_LYSIN" = "#56B4E9"))  

  print(p1)  # Print first page

  # Second Page: Zoomed-In View (Max Y = 5%)
  low_abundance_data <- paired_data %>% filter(Abundance <= 5)

  p2 <- ggplot(low_abundance_data, aes(x = Species, y = Abundance, fill = Species)) +
    geom_boxplot(alpha = 0.7, outlier.shape = NA) +  
    geom_jitter(aes(color = TREATMENT), shape = 16, position = position_jitter(0.2), alpha = 0.6) +  
    geom_line(aes(group = fastqID, color = TREATMENT), alpha = 0.5) +
    scale_x_discrete(labels = c(species, "Salivarius")) +  
    ylim(0, 5) +  # Correctly limit y-axis to 5%
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    labs(title = paste("Zoomed-In: Low Abundance (≤5%) -", species, "vs Salivarius"),
         x = "Bacteria", y = "Relative Abundance (%)", color = "Treatment") +
    scale_fill_viridis_d() +  
    scale_color_manual(values = c("Lysin" = "#E69F00", "NO_LYSIN" = "#56B4E9"))  

  print(p2)  # Print second page
}

# Close PDF file
dev.off()






##########################
## TEST COMPARATIVE BOXPLOT FOR S COMES
##########################


# Load libraries
library(phyloseq)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)

# STEP 1: Subset phyloseq (exclude EB, keep 24 & 48h)
physeq_filtered_NO_EB <- subset_samples(physeq_filtered, TREATMENT != "EB")
physeq_filtered_NO_EB_24_48 <- subset_samples(physeq_filtered_NO_EB, TIME %in% c("24", "48"))

# STEP 2: Transform to relative abundance
physeq_ra <- transform_sample_counts(physeq_filtered_NO_EB_24_48, function(x) x / sum(x))

# STEP 3: Extract OTU, taxonomy, and metadata
otu_df <- as.data.frame(otu_table(physeq_ra))
tax_df <- as.data.frame(tax_table(physeq_ra))
meta_df <- as.data.frame(sample_data(physeq_ra))

# STEP 4: Filter for comes and salivarius
tax_df$ASV <- rownames(tax_df)
filtered_tax <- tax_df %>%
  filter((Genus == "Coprococcus" & Species == "comes") |
         (Genus == "Streptococcus" & Species == "salivarius"))

# STEP 5: Subset OTU table and convert to long format
otu_filtered <- otu_df[rownames(otu_df) %in% filtered_tax$ASV, , drop = FALSE]
otu_filtered$ASV <- rownames(otu_filtered)

otu_long <- otu_filtered %>%
  pivot_longer(-ASV, names_to = "fastqID", values_to = "Abundance") %>%
  left_join(filtered_tax[, c("ASV", "Genus", "Species")], by = "ASV") %>%
  mutate(Taxon = paste(Genus, Species, sep = " "))

# STEP 6: Aggregate to one value per sample per taxon
otu_summary <- otu_long %>%
  group_by(fastqID, Taxon) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")

# STEP 7: Pivot to wide format and remove samples with both taxa absent
otu_wide <- otu_summary %>%
  pivot_wider(names_from = Taxon, values_from = Abundance, values_fill = 0) %>%
  filter(`Coprococcus comes` > 0 | `Streptococcus salivarius` > 0)

# STEP 8: Add pseudocount
pseudo <- 0.0001
otu_wide <- otu_wide %>%
  mutate(
    `Coprococcus comes` = `Coprococcus comes` + pseudo,
    `Streptococcus salivarius` = `Streptococcus salivarius` + pseudo
  )

# STEP 9: Merge with metadata and add SIC_TIME
final_table <- left_join(otu_wide, meta_df, by = "fastqID") %>%
  mutate(SIC_TIME = paste0(SIC_ID, " | ", TIME, "h"))

# STEP 10: Compute direction of change per sample
line_data <- final_table %>%
  mutate(delta = `Coprococcus comes` - `Streptococcus salivarius`,
         direction = ifelse(delta > 0, "Up", "Down")) %>%
  select(fastqID, direction)

# STEP 11: Reshape for plotting and join direction
plot_data <- final_table %>%
  select(fastqID, TREATMENT, SIC_TIME, `Coprococcus comes`, `Streptococcus salivarius`) %>%
  pivot_longer(cols = c(`Coprococcus comes`, `Streptococcus salivarius`),
               names_to = "Taxon", values_to = "Abundance") %>%
  mutate(
    Taxon = factor(Taxon, levels = c("Coprococcus comes", "Streptococcus salivarius")),
    TREATMENT = factor(TREATMENT, levels = c("NO_LYSIN", "Lysin"))
  ) %>%
  left_join(line_data, by = "fastqID")
  
# Adjust Taxon order and SIC_TIME order


plot_data <- plot_data %>%
  mutate(
    Taxon = factor(Taxon, levels = c("Streptococcus salivarius", "Coprococcus comes")),
    SIC_TIME = factor(SIC_TIME, levels = c(
      "SIC8A | 24h", "SIC8A | 48h",
      "SIC9A | 24h", "SIC9A | 48h",
      "SIC10A | 24h", "SIC10A | 48h"
    ))
  )


# Build final plot
p <- ggplot(plot_data, aes(x = Taxon, y = Abundance)) +
  geom_line(aes(group = fastqID, color = direction), alpha = 0.6, linewidth = 0.5) +
  geom_boxplot(aes(fill = Taxon), outlier.shape = NA, alpha = 0.3,
               position = position_dodge(width = 0.75)) +
  geom_point(color = "black",
             position = position_dodge(width = 0.75),
             alpha = 0.5, size = 1.5) +
  scale_color_manual(
    values = c("Up" = "#ff7f0e", "Down" = "#1f77b4"),
    name = "Line direction"
  ) +
  scale_fill_manual(
    values = c(
      "Streptococcus salivarius" = "#1f77b4",
      "Coprococcus comes" = "#ff7f0e"
    )
  ) +
  scale_y_log10() +
  facet_grid(SIC_TIME ~ TREATMENT, switch = "y", scales = "free_y", space = "free") +
  theme_classic(base_size = 11) +
  labs(
    title = "Paired Abundance with Direction-Colored Lines (log10)",
    x = "Taxon",
    y = "Relative Abundance (%) [log10]",
    fill = "Taxon"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text.y.left = element_text(angle = 0, size = 10),
    strip.placement = "outside",
    panel.spacing.y = unit(1.5, "lines"),
    legend.position = "top"
  )

# Show and export
print(p)
ggsave("paired_boxplot_custom_colors_faceted_ordered_log10.pdf", plot = p, width = 12, height = 8)


##################### OTHER GRAPH

# Set factor levels
plot_data <- plot_data %>%
  mutate(
    Taxon = factor(Taxon, levels = c("Streptococcus salivarius", "Coprococcus comes")),
    TREATMENT = factor(TREATMENT, levels = c("NO_LYSIN", "Lysin"))
  )

# Styled and simplified plot
p <- ggplot(plot_data, aes(x = Taxon, y = Abundance)) +
  geom_line(aes(group = fastqID, color = direction), alpha = 0.6, linewidth = 0.5) +
  geom_boxplot(aes(fill = Taxon), outlier.shape = NA, alpha = 0.3,
               position = position_dodge(width = 0.75)) +
  geom_point(color = "black",
             position = position_dodge(width = 0.75),
             alpha = 0.5, size = 1.5) +
  scale_color_manual(
    values = c("Up" = "#ff7f0e", "Down" = "#1f77b4"),
    name = "Line direction"
  ) +
  scale_fill_manual(
    values = c(
      "Streptococcus salivarius" = "#1f77b4",
      "Coprococcus comes" = "#ff7f0e"
    )
  ) +
  scale_y_log10() +
  facet_wrap(~ TREATMENT, nrow = 1) +
  theme_classic(base_size = 11) +
  labs(
    title = "Paired Abundance by Treatment with Direction-Colored Lines (log10)",
    x = "Taxon",
    y = "Relative Abundance (%) [log10]",
    fill = "Taxon"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 10),
    legend.position = "top"
  )

# Print and export
print(p)
ggsave("paired_boxplot_TREATMENT_only_styled.pdf", plot = p, width = 8, height = 5)







##########################
## TEST COMPARATIVE BOXPLOT FOR ANY 2 SPECIES OF CHOICE
##########################


# Load libraries
library(phyloseq)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)

# STEP 1: Subset phyloseq (exclude EB, keep 24 & 48h)
physeq_filtered_NO_EB <- subset_samples(physeq_filtered, TREATMENT != "EB")
physeq_filtered_NO_EB_24_48 <- subset_samples(physeq_filtered_NO_EB, TIME %in% c("24", "48"))

# STEP 2: Transform to relative abundance
physeq_ra <- transform_sample_counts(physeq_filtered_NO_EB_24_48, function(x) x / sum(x))

# STEP 3: Extract OTU, taxonomy, and metadata
otu_df <- as.data.frame(otu_table(physeq_ra))
tax_df <- as.data.frame(tax_table(physeq_ra))
meta_df <- as.data.frame(sample_data(physeq_ra))

# STEP 4: Filter for comes and salivarius
tax_df$ASV <- rownames(tax_df)
filtered_tax <- tax_df %>%
  filter((Genus == "Coprococcus" & Species == "comes") |
         (Genus == "Streptococcus" & Species == "salivarius"))

# STEP 5: Subset OTU table and convert to long format
otu_filtered <- otu_df[rownames(otu_df) %in% filtered_tax$ASV, , drop = FALSE]
otu_filtered$ASV <- rownames(otu_filtered)

otu_long <- otu_filtered %>%
  pivot_longer(-ASV, names_to = "fastqID", values_to = "Abundance") %>%
  left_join(filtered_tax[, c("ASV", "Genus", "Species")], by = "ASV") %>%
  mutate(Taxon = paste(Genus, Species, sep = " "))

# STEP 6: Aggregate to one value per sample per taxon
otu_summary <- otu_long %>%
  group_by(fastqID, Taxon) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")

# STEP 7: Pivot to wide format and remove samples with both taxa absent
otu_wide <- otu_summary %>%
  pivot_wider(names_from = Taxon, values_from = Abundance, values_fill = 0) %>%
  filter(`Coprococcus comes` > 0 | `Streptococcus salivarius` > 0)

# STEP 8: Add pseudocount
pseudo <- 0.0001
otu_wide <- otu_wide %>%
  mutate(
    `Coprococcus comes` = `Coprococcus comes` + pseudo,
    `Streptococcus salivarius` = `Streptococcus salivarius` + pseudo
  )

# STEP 9: Merge with metadata and add SIC_TIME
final_table <- left_join(otu_wide, meta_df, by = "fastqID") %>%
  mutate(SIC_TIME = paste0(SIC_ID, " | ", TIME, "h"))

# STEP 10: Compute direction of change per sample
line_data <- final_table %>%
  mutate(delta = `Coprococcus comes` - `Streptococcus salivarius`,
         direction = ifelse(delta > 0, "Up", "Down")) %>%
  select(fastqID, direction)

# STEP 11: Reshape for plotting and join direction
plot_data <- final_table %>%
  select(fastqID, TREATMENT, SIC_TIME, `Coprococcus comes`, `Streptococcus salivarius`) %>%
  pivot_longer(cols = c(`Coprococcus comes`, `Streptococcus salivarius`),
               names_to = "Taxon", values_to = "Abundance") %>%
  mutate(
    Taxon = factor(Taxon, levels = c("Coprococcus comes", "Streptococcus salivarius")),
    TREATMENT = factor(TREATMENT, levels = c("NO_LYSIN", "Lysin"))
  ) %>%
  left_join(line_data, by = "fastqID")
  
# Adjust Taxon order and SIC_TIME order


plot_data <- plot_data %>%
  mutate(
    Taxon = factor(Taxon, levels = c("Streptococcus salivarius", "Coprococcus comes")),
    SIC_TIME = factor(SIC_TIME, levels = c(
      "SIC8A | 24h", "SIC8A | 48h",
      "SIC9A | 24h", "SIC9A | 48h",
      "SIC10A | 24h", "SIC10A | 48h"
    ))
  )


# Build final plot
p <- ggplot(plot_data, aes(x = Taxon, y = Abundance)) +
  geom_line(aes(group = fastqID, color = direction), alpha = 0.6, linewidth = 0.5) +
  geom_boxplot(aes(fill = Taxon), outlier.shape = NA, alpha = 0.3,
               position = position_dodge(width = 0.75)) +
  geom_point(color = "black",
             position = position_dodge(width = 0.75),
             alpha = 0.5, size = 1.5) +
  scale_color_manual(
    values = c("Up" = "#ff7f0e", "Down" = "#1f77b4"),
    name = "Line direction"
  ) +
  scale_fill_manual(
    values = c(
      "Streptococcus salivarius" = "#1f77b4",
      "Coprococcus comes" = "#ff7f0e"
    )
  ) +
  scale_y_log10() +
  facet_grid(SIC_TIME ~ TREATMENT, switch = "y", scales = "free_y", space = "free") +
  theme_classic(base_size = 11) +
  labs(
    title = "Paired Abundance with Direction-Colored Lines (log10)",
    x = "Taxon",
    y = "Relative Abundance (%) [log10]",
    fill = "Taxon"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text.y.left = element_text(angle = 0, size = 10),
    strip.placement = "outside",
    panel.spacing.y = unit(1.5, "lines"),
    legend.position = "top"
  )

# Show and export
print(p)
ggsave("paired_boxplot_custom_colors_faceted_ordered_log10.pdf", plot = p, width = 12, height = 8)


##################### OTHER GRAPH

# Set factor levels
plot_data <- plot_data %>%
  mutate(
    Taxon = factor(Taxon, levels = c("Streptococcus salivarius", "Coprococcus comes")),
    TREATMENT = factor(TREATMENT, levels = c("NO_LYSIN", "Lysin"))
  )

# Styled and simplified plot
p <- ggplot(plot_data, aes(x = Taxon, y = Abundance)) +
  geom_line(aes(group = fastqID, color = direction), alpha = 0.6, linewidth = 0.5) +
  geom_boxplot(aes(fill = Taxon), outlier.shape = NA, alpha = 0.3,
               position = position_dodge(width = 0.75)) +
  geom_point(color = "black",
             position = position_dodge(width = 0.75),
             alpha = 0.5, size = 1.5) +
  scale_color_manual(
    values = c("Up" = "#ff7f0e", "Down" = "#1f77b4"),
    name = "Line direction"
  ) +
  scale_fill_manual(
    values = c(
      "Streptococcus salivarius" = "#1f77b4",
      "Coprococcus comes" = "#ff7f0e"
    )
  ) +
  scale_y_log10() +
  facet_wrap(~ TREATMENT, nrow = 1) +
  theme_classic(base_size = 11) +
  labs(
    title = "Paired Abundance by Treatment with Direction-Colored Lines (log10)",
    x = "Taxon",
    y = "Relative Abundance (%) [log10]",
    fill = "Taxon"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 10),
    legend.position = "top"
  )

# Print and export
print(p)
ggsave("paired_boxplot_TREATMENT_only_styled.pdf", plot = p, width = 8, height = 5)






##########################
## COMPARATIVE BOXPLOT FOR ANY 2 SPECIES OF CHOICE
##########################

# Load libraries
library(phyloseq)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)


You can :

print(unique(tax_df$Genus)) 
#and
unique(tax_df$Species) to inspect available values.

# OR
# If you want to list the species available for a specific genus of interest, change the genus_of_interest by
# the genus you want to know more about

View(
  tax_df %>%
    filter(Genus == "Bacteroides")
)



# ===== Choose your two species of interest =====
species_1 <- c("Coprococcus", "comes")
species_2 <- c("Streptococcus", "salivarius")

# ===== Create species label for file names =====
species_label_1 <- paste(species_1, collapse = "_")
species_label_2 <- paste(species_2, collapse = "_")
plot_tag <- paste0(species_label_1, "_vs_", species_label_2)

# STEP 1: Subset phyloseq (exclude EB, keep 24 & 48h)
physeq_filtered_NO_EB <- subset_samples(physeq_filtered, TREATMENT != "EB")
physeq_filtered_NO_EB_24_48 <- subset_samples(physeq_filtered_NO_EB, TIME %in% c("24", "48"))

# STEP 2: Transform to relative abundance
physeq_ra <- transform_sample_counts(physeq_filtered_NO_EB_24_48, function(x) x / sum(x))

# STEP 3: Extract OTU, taxonomy, and metadata
otu_df <- as.data.frame(otu_table(physeq_ra))
tax_df <- as.data.frame(tax_table(physeq_ra))
meta_df <- as.data.frame(sample_data(physeq_ra))

# STEP 4: Filter for selected species
tax_df$ASV <- rownames(tax_df)
filtered_tax <- tax_df %>%
  filter((Genus == species_1[1] & Species == species_1[2]) |
         (Genus == species_2[1] & Species == species_2[2]))

# STEP 5: Subset OTU table and convert to long format
otu_filtered <- otu_df[rownames(otu_df) %in% filtered_tax$ASV, , drop = FALSE]
otu_filtered$ASV <- rownames(otu_filtered)

otu_long <- otu_filtered %>%
  pivot_longer(-ASV, names_to = "fastqID", values_to = "Abundance") %>%
  left_join(filtered_tax[, c("ASV", "Genus", "Species")], by = "ASV") %>%
  mutate(Taxon = paste(Genus, Species, sep = " "))

# STEP 6: Aggregate to one value per sample per taxon
otu_summary <- otu_long %>%
  group_by(fastqID, Taxon) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")

# STEP 7: Pivot to wide format and remove samples with both taxa absent
taxon_names <- unique(otu_summary$Taxon)
otu_wide <- otu_summary %>%
  pivot_wider(names_from = Taxon, values_from = Abundance, values_fill = 0) %>%
  filter((.data[[taxon_names[1]]] > 0) | (.data[[taxon_names[2]]] > 0))

# STEP 8: Add pseudocount
pseudo <- 0.0001
otu_wide[[taxon_names[1]]] <- otu_wide[[taxon_names[1]]] + pseudo
otu_wide[[taxon_names[2]]] <- otu_wide[[taxon_names[2]]] + pseudo

# STEP 9: Merge with metadata and add SIC_TIME
final_table <- left_join(otu_wide, meta_df, by = "fastqID") %>%
  mutate(SIC_TIME = paste0(SIC_ID, " | ", TIME, "h"))

# STEP 10: Compute direction of change per sample
line_data <- final_table %>%
  mutate(delta = .data[[taxon_names[1]]] - .data[[taxon_names[2]]],
         direction = ifelse(delta > 0, "Up", "Down")) %>%
  select(fastqID, direction)

# STEP 11: Reshape for plotting and join direction
plot_data <- final_table %>%
  select(fastqID, TREATMENT, SIC_TIME, all_of(taxon_names)) %>%
  pivot_longer(cols = all_of(taxon_names),
               names_to = "Taxon", values_to = "Abundance") %>%
  mutate(
    Taxon = factor(Taxon, levels = taxon_names),
    TREATMENT = factor(TREATMENT, levels = c("NO_LYSIN", "Lysin"))
  ) %>%
  left_join(line_data, by = "fastqID")

# Export underlying data
write_tsv(plot_data, paste0("plot_data_", plot_tag, ".tsv"))

# === Plot 1: Faceted by SIC_TIME & TREATMENT ===
p1 <- ggplot(plot_data, aes(x = Taxon, y = Abundance)) +
  geom_line(aes(group = fastqID, color = direction), alpha = 0.6, linewidth = 0.5) +
  geom_boxplot(aes(fill = Taxon), outlier.shape = NA, alpha = 0.3,
               position = position_dodge(width = 0.75)) +
  geom_point(color = "black",
             position = position_dodge(width = 0.75),
             alpha = 0.5, size = 1.5) +
  scale_color_manual(values = c("Up" = "#ff7f0e", "Down" = "#1f77b4"), name = "Line direction") +
  scale_y_log10() +
  facet_grid(SIC_TIME ~ TREATMENT, switch = "y", scales = "free_y", space = "free") +
  theme_classic(base_size = 11) +
  labs(
    title = paste("Paired Abundance:", species_label_1, "vs", species_label_2),
    x = "Taxon",
    y = "Relative Abundance (%) [log10]",
    fill = "Taxon"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text.y.left = element_text(angle = 0, size = 10),
    strip.placement = "outside",
    panel.spacing.y = unit(1.5, "lines"),
    legend.position = "top"
  )

ggsave(
  paste0("paired_boxplot_faceted_", plot_tag, ".pdf"),
  plot = p1, width = 12, height = 8
)

# === Plot 2: Faceted by TREATMENT only ===
p2 <- ggplot(plot_data, aes(x = Taxon, y = Abundance)) +
  geom_line(aes(group = fastqID, color = direction), alpha = 0.6, linewidth = 0.5) +
  geom_boxplot(aes(fill = Taxon), outlier.shape = NA, alpha = 0.3,
               position = position_dodge(width = 0.75)) +
  geom_point(color = "black",
             position = position_dodge(width = 0.75),
             alpha = 0.5, size = 1.5) +
  scale_color_manual(values = c("Up" = "#ff7f0e", "Down" = "#1f77b4"), name = "Line direction") +
  scale_y_log10() +
  facet_wrap(~ TREATMENT, nrow = 1) +
  theme_classic(base_size = 11) +
  labs(
    title = paste("Abundance by Treatment:", species_label_1, "vs", species_label_2),
    x = "Taxon",
    y = "Relative Abundance (%) [log10]",
    fill = "Taxon"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 10),
    legend.position = "top"
  )

ggsave(
  paste0("paired_boxplot_TREATMENT_only_", plot_tag, ".pdf"),
  plot = p2, width = 8, height = 5
)







###################### EXTRACTING SPECIES LEVEL INFO TO BE USED IN SPIECEASY FOR CO-ABUNDANCE ANALYSES

library(phyloseq)
library(dplyr)
library(tidyr)

# Transform to relative abundance
physeq_species <- tax_glom(physeq, taxrank = "Species")
physeq_ra <- transform_sample_counts(physeq_species, function(x) x / sum(x))

# Get taxonomy and OTU table
tax_df <- as.data.frame(tax_table(physeq_ra))
otu_df <- as.data.frame(otu_table(physeq_ra))

# Append species names to OTU IDs
tax_df$Species_full <- paste(tax_df$Genus, tax_df$Species, sep = " ")
tax_df$Species_full[is.na(tax_df$Species)] <- NA  # Remove malformed names

# Remove NAs
valid_species <- tax_df[!is.na(tax_df$Species_full), ]
otu_filtered <- otu_df[rownames(otu_df) %in% rownames(valid_species), ]
rownames(otu_filtered) <- valid_species$Species_full[match(rownames(otu_filtered), rownames(valid_species))]

# Transpose to samples × species
spiec_input <- t(otu_filtered)

# Save table for verification
write.table(spiec_input, file = "spiec_species_input.tsv", sep = "\t", quote = FALSE)


# Transpose species × samples
fastspar_input <- t(spiec_input)

# Replace NA with pseudocounts
fastspar_input[is.na(fastspar_input)] <- 0.0001

# Set correct header and rownames
fastspar_df <- as.data.frame(fastspar_input)
fastspar_df <- tibble::rownames_to_column(fastspar_df, var = "#OTU ID")

# Write to TSV with BIOM format header
write.table(fastspar_df, file = "fastspar_input.tsv", sep = "\t", quote = FALSE, row.names = FALSE)


